---
title: "ICILS analysis script"
author: "Ismael Aguayo"
date: today
lang: en
fontsize: 14pt
format:
  html:
    code-fold: false
    toc: true
    toc-location: left
    toc-depth: 3
    toc-expand: 1
    toc-title: Contents
    number-sections: true
    number-depth: 3
    theme:
      - cosmo
    code-link: true
    title-block-banner: true
editor_options: 
  chunk_output_type: console
---

# Enviroment setting

```{r}
#install.packages("pacman")
library(pacman)

pacman::p_load(
  dplyr, haven, sjlabelled,  psych,  purrr,  tidyr,  sjPlot,  ggplot2, 
  parameters,  table1,  car,  beeswarm,  lme4, scales, ggrepel, corrplot,
  ggtext, patchwork, lavaan, semTools, DT, knitr, kableExtra, flextable, officer, semTools)

options(scipen = 999)
rm(list = ls()) 

icils <- readRDS(here::here("input", "data", "proc_data", "icils_comparable.rds"))

```

# Pooled model

```{r}
#| warning: false

model_syntax <- '
  gen_dse =~ search_info + source_info + evaluate_info + 
              install_app + share_content + write_text + 
              create_media + insert_image + edit_image + change_settings
  spec_dse =~ develop_webpage + programming + visual_coding
  '

m_pooled <- cfa(model = model_syntax, 
              data = icils,
              estimator = "WLSMV",
              missing = 'pairwise',
              sampling.weights = "TOTWGTS",
              ordered = T)

```

### Fit

```{r}
fit_icils <- fitMeasures(m_pooled,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

knitr::kable(fit_icils, digits = 3)

```

### Item information

```{r}

std_params <- standardizedsolution(m_pooled)


items_table <- std_params %>% 
  filter(op == "=~") %>% 
  select(Factor = lhs, 
         Indicator = rhs, 
         Loading = est.std, 
         SE = se, 
         p_value = pvalue) %>% 
  mutate(R2_Explained = Loading^2)

knitr::kable(items_table, digits = 3, caption = "Loadings and R squared")

```

### Reliability analysis

```{r}
omega_cat <- compRelSEM(m_pooled)

ave_values <- AVE(m_pooled)

reliability_table <- data.frame(
  Factor = names(omega_cat),
  Omega = as.numeric(omega_cat),
  AVE = as.numeric(ave_values)
)

knitr::kable(reliability_table, digits = 3, caption = "Fiabilidad (Omega Categórico) y Validez (AVE)")
```

### Inter-factor correlation

```{r}

std_params %>% 
  filter(op == "~~",      
         lhs != rhs) %>% 
  select(Factor1 = lhs, Factor2 = rhs, Correlation = est.std, 
         SE = se, Z = z, `p-value` = pvalue)

```

# Invariance analysis

## Country invariance

### Model fitting
```{r}
#| warning: false

config_cnt <- cfa(model = model_syntax, 
              data = icils,
              estimator = "WLSMV",
              group = "CNTRY",
              sampling.weights = "TOTWGTS",
              sampling.weights.normalization = "group",
              missing = 'pairwise',
              ordered = T)

metric_cnt <- cfa(model = model_syntax, 
              data = icils,
              estimator = "WLSMV",
              group = "CNTRY",
              sampling.weights = "TOTWGTS",
              sampling.weights.normalization = "group",
              ordered = T,
              missing = 'pairwise',
              group.equal= "loadings")

scalar_cnt <- cfa(model = model_syntax, 
              data = icils,
              estimator = "WLSMV",
              group = "CNTRY",
              sampling.weights = "TOTWGTS",
              sampling.weights.normalization = "group",
              ordered = T,
              missing = 'pairwise',
              group.equal= c("loadings", "intercepts"))

```

### Fit measures estracting

```{r}
fitmeasures_config_cnt <- fitMeasures(config_cnt,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

fitmeasures_metric_cnt <- fitMeasures(metric_cnt,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

fitmeasures_scalar_cnt <- fitMeasures(scalar_cnt,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))
```

### Table processing

```{r}
tabla_completa_cnt <- data.frame(
  Type = c("Country", "", ""),
  Model = c("1. Configural", "2. Metric", "3. Scalar"),
  chisq = c(fitmeasures_config_cnt["chisq"], fitmeasures_metric_cnt["chisq"], fitmeasures_scalar_cnt["chisq"]),
  df = c(fitmeasures_config_cnt["df"], fitmeasures_metric_cnt["df"], fitmeasures_scalar_cnt["df"]),
  CFI = c(fitmeasures_config_cnt["cfi"], fitmeasures_metric_cnt["cfi"], fitmeasures_scalar_cnt["cfi"]),
  TLI = c(fitmeasures_config_cnt["tli"], fitmeasures_metric_cnt["tli"], fitmeasures_scalar_cnt["tli"]),
  RMSEA = c(fitmeasures_config_cnt["rmsea"], fitmeasures_metric_cnt["rmsea"], fitmeasures_scalar_cnt["rmsea"])
)

# Deltas

tabla_completa_cnt$delta_chisq <- c(NA, diff(tabla_completa_cnt$chisq))
tabla_completa_cnt$delta_df    <- c(NA, diff(tabla_completa_cnt$df))
tabla_completa_cnt$delta_CFI   <- c(NA, diff(tabla_completa_cnt$CFI))
tabla_completa_cnt$delta_RMSEA <- c(NA, diff(tabla_completa_cnt$RMSEA))

# Deltas display

tabla_completa_cnt$delta_CFI_display <- sapply(1:nrow(tabla_completa_cnt), function(i) {
  if (i == 1) return(NA)
  
  if (tabla_completa_cnt$delta_CFI[i] >= -0.004) {
    return("< -0.004")
  } else {
    return(sprintf("%.3f", tabla_completa_cnt$delta_CFI[i]))
  }
})

tabla_completa_cnt$delta_RMSEA_display <- sapply(1:nrow(tabla_completa_cnt), function(i) {
  if (i == 1) return(NA)
  
  umbral <- if (i == 2) 0.050 else 0.010
  
  if (tabla_completa_cnt$delta_RMSEA[i] <= umbral) {
    return(paste0("< ", umbral))
  } else {
    return(sprintf("%.3f", tabla_completa_cnt$delta_RMSEA[i]))
  }
})

# ANOVA for significance

anova_cnt <- anova(config_cnt, metric_cnt, scalar_cnt)

tabla_completa_cnt$p_value_delta_chisq <- case_when(anova_cnt$`Pr(>Chisq)`< 0.01 ~ "< 0.01",
                                                  anova_cnt$`Pr(>Chisq)`< 0.05 ~ "< 0.05",
                                                  anova_cnt$`Pr(>Chisq)`> 0.05 ~ sprintf("%.2f", anova_cnt$`Pr(>Chisq)`))
tabla_completa_cnt$Decision <- c(
  NA,  # Configural
  # Metric: ΔCFI ≥ -0.004 Y ΔRMSEA ≤ 0.050
  ifelse(tabla_completa_cnt$delta_CFI[2] >= -0.004 & 
         tabla_completa_cnt$delta_RMSEA[2] <= 0.050, "Yes", "No"),
  # Scalar: ΔCFI ≥ -0.004 Y ΔRMSEA ≤ 0.010
  ifelse(tabla_completa_cnt$delta_CFI[3] >= -0.004 & 
         tabla_completa_cnt$delta_RMSEA[3] <= 0.010, "Yes", "No")
)
```

## Gender invariance

### Model fitting
```{r}
#| warning: false

config_sex <- cfa(model = model_syntax, 
              data = icils,
              estimator = "WLSMV",
              group = "sex",
              sampling.weights = "TOTWGTS",
              sampling.weights.normalization = "group",
              missing = 'pairwise',
              ordered = T)

metric_sex <- cfa(model = model_syntax, 
              data = icils,
              estimator = "WLSMV",
              group = "sex",
              sampling.weights = "TOTWGTS",
              sampling.weights.normalization = "group",
              ordered = T,
              missing = 'pairwise',
              group.equal= "loadings")

scalar_sex <- cfa(model = model_syntax, 
              data = icils,
              estimator = "WLSMV",
              group = "sex",
              sampling.weights = "TOTWGTS",
              sampling.weights.normalization = "group",
              ordered = T,
              missing = 'pairwise',
              group.equal= c("loadings", "intercepts"))

```

### Fit measures stracting

```{r}
fitmeasures_config_sex <- fitMeasures(config_sex,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

fitmeasures_metric_sex <- fitMeasures(metric_sex,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

fitmeasures_scalar_sex <- fitMeasures(scalar_sex,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))
```

### Table processing
```{r}
tabla_completa_sex <- data.frame(
  Type = c("Gender", "", ""),
  Model = c("1. Configural", "2. Metric", "3. Scalar"),
  chisq = c(fitmeasures_config_sex["chisq"], fitmeasures_metric_sex["chisq"], fitmeasures_scalar_sex["chisq"]),
  df = c(fitmeasures_config_sex["df"], fitmeasures_metric_sex["df"], fitmeasures_scalar_sex["df"]),
  CFI = c(fitmeasures_config_sex["cfi"], fitmeasures_metric_sex["cfi"], fitmeasures_scalar_sex["cfi"]),
  TLI = c(fitmeasures_config_sex["tli"], fitmeasures_metric_sex["tli"], fitmeasures_scalar_sex["tli"]),
  RMSEA = c(fitmeasures_config_sex["rmsea"], fitmeasures_metric_sex["rmsea"], fitmeasures_scalar_sex["rmsea"])
)

# Deltas

tabla_completa_sex$delta_chisq <- c(NA, diff(tabla_completa_sex$chisq))
tabla_completa_sex$delta_df    <- c(NA, diff(tabla_completa_sex$df))
tabla_completa_sex$delta_CFI   <- c(NA, diff(tabla_completa_sex$CFI))
tabla_completa_sex$delta_RMSEA <- c(NA, diff(tabla_completa_sex$RMSEA))

# Deltas display

tabla_completa_sex$delta_CFI_display <- sapply(1:nrow(tabla_completa_sex), function(i) {
  if (i == 1) return(NA)
  
  if (tabla_completa_sex$delta_CFI[i] >= -0.004) {
    return("< -0.004")
  } else {
    return(sprintf("%.3f", tabla_completa_sex$delta_CFI[i]))
  }
})

tabla_completa_sex$delta_RMSEA_display <- sapply(1:nrow(tabla_completa_sex), function(i) {
  if (i == 1) return(NA)
  
  umbral <- if (i == 2) 0.050 else 0.010
  
  if (tabla_completa_sex$delta_RMSEA[i] <= umbral) {
    return(paste0("< ", umbral))
  } else {
    return(sprintf("%.3f", tabla_completa_sex$delta_RMSEA[i]))
  }
})

# ANOVA for significance

anova_sex <- anova(config_sex, metric_sex, scalar_sex)

tabla_completa_sex$p_value_delta_chisq <- case_when(anova_sex$`Pr(>Chisq)`< 0.01 ~ "< 0.01",
                                                  anova_sex$`Pr(>Chisq)`< 0.05 ~ "< 0.05",
                                                  anova_sex$`Pr(>Chisq)`> 0.05 ~ sprintf("%.2f", anova_sex$`Pr(>Chisq)`))
tabla_completa_sex$Decision <- c(
  NA,  # Configural
  # Metric: ΔCFI ≥ -0.004 Y ΔRMSEA ≤ 0.050
  ifelse(tabla_completa_sex$delta_CFI[2] >= -0.004 & 
         tabla_completa_sex$delta_RMSEA[2] <= 0.050, "Yes", "No"),
  # Scalar: ΔCFI ≥ -0.004 Y ΔRMSEA ≤ 0.010
  ifelse(tabla_completa_sex$delta_CFI[3] >= -0.004 & 
         tabla_completa_sex$delta_RMSEA[3] <= 0.010, "Yes", "No")
)
```

## Final table formatting
```{r}
tabla_completa_cnt_export <- tabla_completa_cnt %>%
  select(Type, Model, chisq, df, CFI, TLI, RMSEA,
         delta_chisq, delta_df, 
         delta_CFI_display, delta_RMSEA_display,
         p_value_delta_chisq, Decision) %>%
  rename(delta_CFI = delta_CFI_display,
         delta_RMSEA = delta_RMSEA_display)

tabla_completa_sex_export <- tabla_completa_sex %>%
  select(Type, Model, chisq, df, CFI, TLI, RMSEA,
         delta_chisq, delta_df, 
         delta_CFI_display, delta_RMSEA_display,
         p_value_delta_chisq, Decision) %>%
  rename(delta_CFI = delta_CFI_display,
         delta_RMSEA = delta_RMSEA_display)

tabla_unificada_icils <- bind_rows(tabla_completa_cnt_export, tabla_completa_sex_export)

```

```{r}

colnames(tabla_unificada_icils) <- c(
  "Type",
  "Model", 
  "\u03C7\u00B2",      # χ²
  "df",
  "CFI",
  "TLI",
  "RMSEA",
  "\u0394\u03C7\u00B2", # Δχ²
  "\u0394df",          # Δdf
  "\u0394CFI",         # ΔCFI
  "\u0394RMSEA",       # ΔRMSEA
  "p",
  "Decision"
)

ft_icils <- flextable(tabla_unificada_icils) %>%
  colformat_double(j = c("\u03C7\u00B2", "df", "\u0394\u03C7\u00B2", "\u0394df"), digits = 0) %>%
  colformat_double(j = c("CFI", "TLI", "RMSEA", "\u0394CFI", "\u0394RMSEA", "p"), digits = 3) %>%  border_remove() %>%
  hline_top(border = officer::fp_border(width = 2), part = "all") %>%
  hline_bottom(border = officer::fp_border(width = 2), part = "all") %>%
  hline(i = 1, border = officer::fp_border(width = 1.5), part = "header") %>%
  bold(part = "header") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 9, part = "body") %>%
  fontsize(size = 10, part = "header") %>%
  padding(padding = 3, part = "all") %>%
  align(align = "center", part = "header") %>%
  align(align = "left", part = "body", j = 1:2) %>%
  align(align = "center", part = "body", j = 3:ncol(tabla_unificada_icils)) %>%
  autofit() %>%
  fit_to_width(max_width = 7.5)

# Which cols to color
delta_cfi_col <- which(colnames(tabla_unificada_icils) == "\u0394CFI")
delta_rmsea_col <- which(colnames(tabla_unificada_icils) == "\u0394RMSEA")
decision_col <- which(colnames(tabla_unificada_icils) == "Decision")
# Which rows to color
rows_cfi_fail <- which(!grepl("<", tabla_unificada_icils[[delta_cfi_col]]) & 
                       !is.na(tabla_unificada_icils[[delta_cfi_col]]))
rows_rmsea_fail <- which(!grepl("<", tabla_unificada_icils[[delta_rmsea_col]]) & 
                         !is.na(tabla_unificada_icils[[delta_rmsea_col]]))
rows_decision_no <- which(tabla_unificada_icils$Decision == "No")
# Red when dont achieve invariance
if(length(rows_cfi_fail) > 0) {
  ft_icils <- ft_icils %>%
    bg(i = rows_cfi_fail, j = delta_cfi_col, bg = "#FFCCCB") %>%
    color(i = rows_cfi_fail, j = delta_cfi_col, color = "#8B0000")
}

if(length(rows_rmsea_fail) > 0) {
  ft_icils <- ft_icils %>%
    bg(i = rows_rmsea_fail, j = delta_rmsea_col, bg = "#FFCCCB") %>%
    color(i = rows_rmsea_fail, j = delta_rmsea_col, color = "#8B0000")
}

if(length(rows_decision_no) > 0) {
  ft_icils <- ft_icils %>%
    bg(i = rows_decision_no, j = decision_col, bg = "#FFCCCB") %>%
    color(i = rows_decision_no, j = decision_col, color = "#8B0000") %>%
    bold(i = rows_decision_no, j = decision_col)
}

ft_icils
```

```{r}
#| eval: false

# Save
save_as_docx(ft_icils, path = "output/figures/tables/invariance_table_icils.docx")
save_as_image(ft_icils, path = "output/figures/tables/invariance_table_icils.png")
save_as_image(ft_icils, path = "output/article/images/invariance_table_icils.png")
saveRDS(ft_icils, "./output/article/tables/invariance_table_icils.rds")

```

# Loadings and chi analysis (appendix)

```{r}

all_params <- parameterEstimates(scalar_cnt, standardized = TRUE)

factor_loadings_country <- all_params %>%
  filter(op == "=~")

test_stats <- lavInspect(scalar_cnt, "test")
chisq_values <- test_stats$scaled.shifted$stat.group
country_names_lavaan <- attr(test_stats, "info")$group.label

chisq_map <- data.frame(
  CNTRY = country_names_lavaan,
  Chisqr = as.numeric(chisq_values)
)

label_dictionary <- icils %>%
  select(CNTRY) %>%
  distinct() %>%
  mutate(
    label_text = as.character(haven::as_factor(CNTRY)) 
  )

n_countries <- icils |> 
  group_by(CNTRY) |> 
  select(CNTRY) |> 
  mutate(N = n()) |> 
  distinct(CNTRY, .keep_all = TRUE)

groups_code <- lavInspect(scalar_cnt, "group.label")

matched_labels <- label_dictionary$label_text[match(groups_code, label_dictionary$CNTRY)]

country_map <- data.frame(
  group = 1:length(groups_code),
  CNTRY = groups_code,
  Country = matched_labels
)

factor_loadings_final <- factor_loadings_country %>%
  left_join(country_map, by = "group") %>%
  left_join(n_countries, by = "CNTRY") %>%
  left_join(chisq_map, by = "CNTRY") %>%
  select(Country, Item = rhs, Factor = lhs, Loading = std.all, N, Chisqr)


```

## Table for appendix

```{r}
loadings_table <- factor_loadings_final %>%
  select(Country, Factor, Item, Loading, N, Chisqr) %>%
  arrange(desc(Chisqr), Factor)

borde_grueso <- fp_border(width = 2)
borde_medio  <- fp_border(width = 1.5)
borde_gris   <- fp_border(color = "grey80", width = 0.5)

ft_loadings <- flextable(loadings_table) %>%
  
  set_header_labels(
    Country = "Country",
    Factor  = "Latent Factor",
    Item    = "Item",
    Loading = "Std. Loading",
    N       = "N",
    Chisqr  = "\u03C7\u00B2"
  ) %>%
  
  merge_v(j = c("Country", "Factor", "N", "Chisqr")) %>%
  
  colformat_double(j = "Loading", digits = 3) %>%
  colformat_double(j = "Chisqr", digits = 2) %>%

  colformat_double(j = "N", digits = 0) %>% 
  
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 9, part = "body") %>%
  fontsize(size = 10, part = "header") %>%
  bold(part = "header") %>%
  
  padding(padding = 3, part = "all") %>%
  align(align = "center", part = "header") %>%
  
  align(align = "center", j = c("Factor", "Loading", "N", "Chisqr"), part = "body") %>% 
  align(align = "left", j = c("Country", "Item"), part = "body") %>%
  
  valign(valign = "center", j = c("Country", "Factor", "N", "Chisqr"), part = "body") %>%

  border_remove() %>%
  
  hline(i = ~ !duplicated(Country, fromLast = TRUE), 
        border = borde_gris) %>%
  
  hline_top(border = borde_grueso, part = "all") %>%
  hline_bottom(border = borde_grueso, part = "all") %>%
  hline(part = "header", border = borde_medio) %>%
  
  autofit() %>%
  fit_to_width(max_width = 7.5)

ft_loadings
```

### Table for report

```{r}

datatable(
  factor_loadings_final, 
  rownames = FALSE,
  filter = 'top',
  options = list(
    pageLength = 15,
    autoWidth = TRUE
  ),
  caption = 'Interactive table: Loadings and Chi square by country'
) %>%
  formatRound(columns = 'Loading', digits = 3) %>% 
  formatRound(columns = 'Chisqr', digits = 2)

```

```{r}
#| eval: false

# Saving table
saveRDS(ft_loadings, "./output/article/tables/loadings_chi_icils.rds")
save_as_docx(ft_loadings, path = "./output/figures/tables/loadings_chi_icils.docx")
save_as_image(ft_loadings, path = "./output/figures/tables/loadings_chi_icils.png")
```