---
title: "ICILS analysis script"
author: "Ismael Aguayo"
date: today
lang: en
fontsize: 14pt
format:
  html:
    code-fold: false
    toc: true
    toc-location: left
    toc-depth: 3
    toc-expand: 1
    toc-title: Contents
    number-sections: true
    number-depth: 3
    theme:
      - cosmo
    code-link: true
    title-block-banner: true
editor_options: 
  chunk_output_type: console
---

# Enviroment setting

```{r}
#install.packages("pacman")
library(pacman)

pacman::p_load(
  dplyr, haven, sjlabelled,  psych,  purrr,  tidyr,  sjPlot,  ggplot2, 
  parameters,  table1,  car,  beeswarm,  lme4, scales, ggrepel, corrplot,
  ggtext, patchwork, lavaan, semTools, DT, knitr, kableExtra, flextable, officer, ggbeeswarm, stringr)

options(scipen = 999)
rm(list = ls()) 

icils <- readRDS(here::here("input", "data", "proc_data", "icils_comparable.rds"))

```

# Pooled model

```{r}
#| warning: false

model_syntax <- '
  gen_dse =~ search_info + source_info + evaluate_info + 
              install_app + share_content + write_text + 
              create_media + insert_image + edit_image + change_settings
  spec_dse =~ develop_webpage + programming + visual_coding
  '

m_pooled <- cfa(model = model_syntax, 
              data = icils,
              estimator = "WLSMV",
              missing = 'pairwise',
              sampling.weights = "SENWT",
              ordered = T)

```

## Fit

```{r}
fit_icils <- fitMeasures(m_pooled,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

knitr::kable(fit_icils, digits = 3)

```

The initial model shows a sub-optimal fit to the data, with a RMSEA above the acceptable threshold of 0.08 ($RMSEA = .088$). Therefore, we proceed to examine modification indices and residual covariances to identify potential areas for model improvement.

## Modification indices and residual covariances

```{r}
mod_indices_1 <- modindices(m_pooled, sort = TRUE, maximum.number = 5)

knitr::kable(mod_indices_1, digits = 3)

matriz_residuales_1 <- resid(m_pooled, type = "cor")$cov

corrplot(matriz_residuales_1, 
         is.corr = FALSE,       
         method = "color", 
         tl.col = "black",      
         cl.pos = "r",          
         type = "lower",
         title = "Lavaan residual correlation matrix")

```

The source info item shows a strong modification index (mi = 13,668.28) because of its cross-loading on the specialized DSE factor. Adittionally, the indicator has strong possitive residual correlations with other items from the specialized DSE factor, such as programming and visual coding. Therefore, we proceed to re-specify the model by removing the source info item from the general DSE factor.

# Re-specification
```{r}
#| warning: false

model_syntax_2 <- '
  gen_dse =~ search_info + evaluate_info + 
              install_app + share_content + write_text + 
              create_media + insert_image + edit_image + change_settings
  spec_dse =~ develop_webpage + programming + visual_coding
  '

m_pooled_2 <- cfa(model = model_syntax_2, 
              data = icils,
              estimator = "WLSMV",
              missing = 'pairwise',
              sampling.weights = "SENWT",
              ordered = T)

```

## Fit

```{r}
fit_icils <- fitMeasures(m_pooled_2,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

knitr::kable(fit_icils, digits = 3)

```

The re-specified model shows an improved fit to the data, with a RMSEA below the acceptable threshold of 0.08 ($RMSEA = .077$).

## Modification indices and residual covariances

```{r}
mod_indices_2 <- modindices(m_pooled_2, sort = TRUE, maximum.number = 5)

knitr::kable(mod_indices_2, digits = 3)

matriz_residuales_2 <- resid(m_pooled_2, type = "cor")$cov

corrplot(matriz_residuales_2, 
         is.corr = FALSE,       
         method = "color", 
         tl.col = "black",      
         cl.pos = "r",          
         type = "lower",
         title = "Lavaan residual correlation matrix")

```

Develop webpage shows a strong modification index (mi = 4,707.51) because of its cross-loading on the general DSE factor. However, due to constraints regarding the number of indicators per factor, this could not be removed from the specialized DSE factor. Create media also shows a notable modification index (mi = 3,982.96) due to its cross-loading on the specialized DSE factor, and strong possitive residual correlations with develop webpage. To mantain the parsimony of the model, this item was also removed.    

# Re-specification 2
```{r}
#| warning: false

model_syntax_3 <- '
  gen_dse =~ search_info + evaluate_info + 
              install_app + share_content + write_text + 
              + insert_image + edit_image + change_settings
  spec_dse =~ develop_webpage + programming + visual_coding
  '

m_pooled_3 <- cfa(model = model_syntax_3, 
              data = icils,
              estimator = "WLSMV",
              missing = 'pairwise',
              sampling.weights = "SENWT",
              ordered = T)

```

## Fit

```{r}
fit_icils <- fitMeasures(m_pooled_3,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

knitr::kable(fit_icils, digits = 3)

```

The re-specified model shows an improved fit to the data, with with key fit indices above commonly accepted thresholds ($CFI = .979$; $TLI = .973$, $RMSEA = .075$). 

## Modindices and residuals details for appendix

### Modindices table
```{r}
mod_indices_table <- bind_rows(
  mod_indices_1 %>% 
    mutate(Stage = "First Re-specification"),
  mod_indices_2 %>% 
    mutate(Stage = "Second Re-specification")
) %>%
  unite("Parameter", lhs, op, rhs, sep = " ", remove = TRUE) %>%
  select(Stage, Parameter, mi, epc, sepc.all) %>%

  mutate(Parameter = str_replace_all(Parameter, "=~", "=\u223C")) %>% 
  mutate(Parameter = str_replace_all(Parameter, "~~", "\u223C\u223C"))

head(mod_indices_table)
```

```{r}
borde_grueso <- fp_border(width = 2)
borde_medio  <- fp_border(width = 1.5)
borde_gris   <- fp_border(color = "grey80", width = 0.5)

ft_mi <- flextable(mod_indices_table) %>%
  
  set_header_labels(
    Stage     = "Analysis Stage",
    Parameter = "Suggested Modification",
    mi        = "MI",
    epc       = "EPC",
    sepc.all  = "Std. EPC"
  ) %>%
  
  add_footer_lines(
    values = "Note. \u223C\u223C indicates correlation; =\u223C indicates cross-loading."
  ) %>%
  
  merge_v(j = "Stage") %>% 
  colformat_double(j = c("mi", "epc", "sepc.all"), digits = 3) %>%
  

  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 9, part = "body") %>%
  fontsize(size = 10, part = "header") %>%
  fontsize(size = 9, part = "footer") %>% 
  italic(part = "footer") %>%
  
  bold(part = "header") %>%
  
  padding(padding = 3, part = "all") %>%
  align(align = "center", part = "header") %>%
  align(align = "center", j = c("mi", "epc", "sepc.all"), part = "body") %>%
  align(align = "left", j = c("Parameter"), part = "body") %>%
  valign(valign = "center", j = "Stage", part = "body") %>% 
  
  border_remove() %>%
  hline(i = ~ !duplicated(Stage, fromLast = TRUE), border = borde_gris) %>%
  hline_top(border = borde_grueso, part = "all") %>%
  hline_bottom(border = borde_grueso, part = "body") %>% 
  hline(part = "header", border = borde_medio) %>%
  
  autofit() %>%
  fit_to_width(max_width = 7.5)

ft_mi
```

#### Saving table for article

```{r}
#| eval: false

# Saving table
saveRDS(ft_mi, "./output/article/tables/mod_indices_icils.rds")
save_as_docx(ft_mi, path = "./output/figures/tables/mod_indices_icils.docx")
save_as_image(ft_mi, path = "./output/figures/tables/mod_indices_icils.png")
```

### Residuals table

```{r}

library(flextable)

get_upper_tri <- function(matriz, model_name) {
  matriz[lower.tri(matriz, diag = TRUE)] <- NA
  as.data.frame(as.table(matriz)) %>%
    na.omit() %>%
    rename(Item1 = Var1, Item2 = Var2, Residual = Freq) %>%
    mutate(Model = model_name)
}

df_res_1 <- get_upper_tri(matriz_residuales_1, "Model 1 (Initial)")
df_res_2 <- get_upper_tri(matriz_residuales_2, "Model 2 (Re-specified)")

df_plot <- bind_rows(df_res_1, df_res_2)

cutoff <- 0.1 

table_residuals <- df_plot %>% 
  filter(abs(Residual) > cutoff) %>%
  arrange(Model, desc(abs(Residual))) %>%
  unite("Pair", Item1, Item2, sep = " \u223C\u223C ")

ft_resid <- flextable(table_residuals) %>%
  set_header_labels(
    Model = "Model Stage",
    Pair = "Item Pair",
    Residual = "Residual Corr."
  ) %>%

  merge_v(j = "Model") %>%
  colformat_double(j = "Residual", digits = 3) %>%
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 10, part = "all") %>%
  bold(part = "header") %>%

  border_remove() %>%
  hline_top(border = fp_border(width = 2), part = "all") %>%
  hline_bottom(border = fp_border(width = 2), part = "body") %>%
  hline(part = "header", border = fp_border(width = 1.5)) %>%
  hline(i = ~ !duplicated(Model, fromLast = TRUE), border = fp_border(color = "grey", width = 0.5)) %>%
  
  autofit() %>%
  add_footer_lines(values = paste("Note. Only pairs with residual correlations >", cutoff, "are shown.")) %>%
  fontsize(size = 9, part = "footer") %>% 
  italic(part = "footer")

ft_resid

```

#### Saving for article

```{r}
#| eval: false

# Saving table
saveRDS(ft_resid, "./output/article/tables/residuals_icils.rds")
save_as_docx(ft_resid, path = "./output/figures/tables/residuals_icils.docx")
save_as_image(ft_resid, path = "./output/figures/tables/residuals_icils.png")
```

## Item information

```{r}

std_params <- standardizedsolution(m_pooled_3)


items_table <- std_params %>% 
  filter(op == "=~") %>% 
  select(Factor = lhs, 
         Indicator = rhs, 
         Loading = est.std, 
         SE = se, 
         p_value = pvalue) %>% 
  mutate(R2_Explained = Loading^2)

knitr::kable(items_table, digits = 3, caption = "Loadings and R squared")

```

## Reliability analysis

```{r}
omega_cat <- compRelSEM(m_pooled_3)

ave_values <- AVE(m_pooled_3)

reliability_table <- data.frame(
  Factor = names(omega_cat),
  Omega = as.numeric(omega_cat),
  AVE = as.numeric(ave_values)
)

knitr::kable(reliability_table, digits = 3, caption = "Fiabilidad (Omega Categórico) y Validez (AVE)")
```

### Inter-factor correlation

```{r}

std_params %>% 
  filter(op == "~~",      
         lhs != rhs) %>% 
  select(Factor1 = lhs, Factor2 = rhs, Correlation = est.std, 
         SE = se, Z = z, `p-value` = pvalue)

```

# Invariance analysis

## Country invariance

### Model fitting
```{r}
#| warning: false

config_cnt <- cfa(model = model_syntax_3, 
              data = icils,
              estimator = "WLSMV",
              group = "CNTRY",
              sampling.weights = "SENWT",
              sampling.weights.normalization = "group",
              missing = 'pairwise',
              ordered = T)

metric_cnt <- cfa(model = model_syntax_3, 
              data = icils,
              estimator = "WLSMV",
              group = "CNTRY",
              sampling.weights = "SENWT",
              sampling.weights.normalization = "group",
              ordered = T,
              missing = 'pairwise',
              group.equal= "loadings")

scalar_cnt <- cfa(model = model_syntax_3, 
              data = icils,
              estimator = "WLSMV",
              group = "CNTRY",
              sampling.weights = "SENWT",
              sampling.weights.normalization = "group",
              ordered = T,
              missing = 'pairwise',
              group.equal= c("loadings", "intercepts"))

```

### Fit measures estracting

```{r}
fitmeasures_config_cnt <- fitMeasures(config_cnt,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

fitmeasures_metric_cnt <- fitMeasures(metric_cnt,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

fitmeasures_scalar_cnt <- fitMeasures(scalar_cnt,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))
```

### Table processing

```{r}
tabla_completa_cnt <- data.frame(
  Type = c("Country", "", ""),
  Model = c("1. Configural", "2. Metric", "3. Scalar"),
  chisq = c(fitmeasures_config_cnt["chisq"], fitmeasures_metric_cnt["chisq"], fitmeasures_scalar_cnt["chisq"]),
  df = c(fitmeasures_config_cnt["df"], fitmeasures_metric_cnt["df"], fitmeasures_scalar_cnt["df"]),
  CFI = c(fitmeasures_config_cnt["cfi"], fitmeasures_metric_cnt["cfi"], fitmeasures_scalar_cnt["cfi"]),
  TLI = c(fitmeasures_config_cnt["tli"], fitmeasures_metric_cnt["tli"], fitmeasures_scalar_cnt["tli"]),
  RMSEA = c(fitmeasures_config_cnt["rmsea"], fitmeasures_metric_cnt["rmsea"], fitmeasures_scalar_cnt["rmsea"])
)

# Deltas

tabla_completa_cnt$delta_chisq <- c(NA, diff(tabla_completa_cnt$chisq))
tabla_completa_cnt$delta_df    <- c(NA, diff(tabla_completa_cnt$df))
tabla_completa_cnt$delta_CFI   <- c(NA, diff(tabla_completa_cnt$CFI))
tabla_completa_cnt$delta_RMSEA <- c(NA, diff(tabla_completa_cnt$RMSEA))

# Deltas display

tabla_completa_cnt$delta_CFI_display <- sapply(1:nrow(tabla_completa_cnt), function(i) {
  if (i == 1) return(NA)
  
  if (tabla_completa_cnt$delta_CFI[i] >= -0.004) {
    return("< -0.004")
  } else {
    return(sprintf("%.3f", tabla_completa_cnt$delta_CFI[i]))
  }
})

tabla_completa_cnt$delta_RMSEA_display <- sapply(1:nrow(tabla_completa_cnt), function(i) {
  if (i == 1) return(NA)
  
  umbral <- if (i == 2) 0.050 else 0.010
  
  if (tabla_completa_cnt$delta_RMSEA[i] <= umbral) {
    return(paste0("< ", umbral))
  } else {
    return(sprintf("%.3f", tabla_completa_cnt$delta_RMSEA[i]))
  }
})

# ANOVA for significance

anova_cnt <- anova(config_cnt, metric_cnt, scalar_cnt)

tabla_completa_cnt$p_value_delta_chisq <- case_when(anova_cnt$`Pr(>Chisq)`< 0.01 ~ "< 0.01",
                                                  anova_cnt$`Pr(>Chisq)`< 0.05 ~ "< 0.05",
                                                  anova_cnt$`Pr(>Chisq)`> 0.05 ~ sprintf("%.2f", anova_cnt$`Pr(>Chisq)`))
tabla_completa_cnt$Decision <- c(
  NA,  # Configural
  # Metric: ΔCFI ≥ -0.004 Y ΔRMSEA ≤ 0.050
  ifelse(tabla_completa_cnt$delta_CFI[2] >= -0.004 & 
         tabla_completa_cnt$delta_RMSEA[2] <= 0.050, "Yes", "No"),
  # Scalar: ΔCFI ≥ -0.004 Y ΔRMSEA ≤ 0.010
  ifelse(tabla_completa_cnt$delta_CFI[3] >= -0.004 & 
         tabla_completa_cnt$delta_RMSEA[3] <= 0.010, "Yes", "No")
)
```

## Gender invariance

### Model fitting
```{r}
#| warning: false

config_sex <- cfa(model = model_syntax_3, 
              data = icils,
              estimator = "WLSMV",
              group = "sex",
              sampling.weights = "SENWT",
              sampling.weights.normalization = "group",
              missing = 'pairwise',
              ordered = T)

metric_sex <- cfa(model = model_syntax_3, 
              data = icils,
              estimator = "WLSMV",
              group = "sex",
              sampling.weights = "SENWT",
              sampling.weights.normalization = "group",
              ordered = T,
              missing = 'pairwise',
              group.equal= "loadings")

scalar_sex <- cfa(model = model_syntax_3, 
              data = icils,
              estimator = "WLSMV",
              group = "sex",
              sampling.weights = "SENWT",
              sampling.weights.normalization = "group",
              ordered = T,
              missing = 'pairwise',
              group.equal= c("loadings", "intercepts"))

```

### Fit measures stracting

```{r}
fitmeasures_config_sex <- fitMeasures(config_sex,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

fitmeasures_metric_sex <- fitMeasures(metric_sex,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

fitmeasures_scalar_sex <- fitMeasures(scalar_sex,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))
```

### Table processing
```{r}
tabla_completa_sex <- data.frame(
  Type = c("Gender", "", ""),
  Model = c("1. Configural", "2. Metric", "3. Scalar"),
  chisq = c(fitmeasures_config_sex["chisq"], fitmeasures_metric_sex["chisq"], fitmeasures_scalar_sex["chisq"]),
  df = c(fitmeasures_config_sex["df"], fitmeasures_metric_sex["df"], fitmeasures_scalar_sex["df"]),
  CFI = c(fitmeasures_config_sex["cfi"], fitmeasures_metric_sex["cfi"], fitmeasures_scalar_sex["cfi"]),
  TLI = c(fitmeasures_config_sex["tli"], fitmeasures_metric_sex["tli"], fitmeasures_scalar_sex["tli"]),
  RMSEA = c(fitmeasures_config_sex["rmsea"], fitmeasures_metric_sex["rmsea"], fitmeasures_scalar_sex["rmsea"])
)

# Deltas

tabla_completa_sex$delta_chisq <- c(NA, diff(tabla_completa_sex$chisq))
tabla_completa_sex$delta_df    <- c(NA, diff(tabla_completa_sex$df))
tabla_completa_sex$delta_CFI   <- c(NA, diff(tabla_completa_sex$CFI))
tabla_completa_sex$delta_RMSEA <- c(NA, diff(tabla_completa_sex$RMSEA))

# Deltas display

tabla_completa_sex$delta_CFI_display <- sapply(1:nrow(tabla_completa_sex), function(i) {
  if (i == 1) return(NA)
  
  if (tabla_completa_sex$delta_CFI[i] >= -0.004) {
    return("< -0.004")
  } else {
    return(sprintf("%.3f", tabla_completa_sex$delta_CFI[i]))
  }
})

tabla_completa_sex$delta_RMSEA_display <- sapply(1:nrow(tabla_completa_sex), function(i) {
  if (i == 1) return(NA)
  
  umbral <- if (i == 2) 0.050 else 0.010
  
  if (tabla_completa_sex$delta_RMSEA[i] <= umbral) {
    return(paste0("< ", umbral))
  } else {
    return(sprintf("%.3f", tabla_completa_sex$delta_RMSEA[i]))
  }
})

# ANOVA for significance

anova_sex <- anova(config_sex, metric_sex, scalar_sex)

tabla_completa_sex$p_value_delta_chisq <- case_when(anova_sex$`Pr(>Chisq)`< 0.01 ~ "< 0.01",
                                                  anova_sex$`Pr(>Chisq)`< 0.05 ~ "< 0.05",
                                                  anova_sex$`Pr(>Chisq)`> 0.05 ~ sprintf("%.2f", anova_sex$`Pr(>Chisq)`))
tabla_completa_sex$Decision <- c(
  NA,  # Configural
  # Metric: ΔCFI ≥ -0.004 Y ΔRMSEA ≤ 0.050
  ifelse(tabla_completa_sex$delta_CFI[2] >= -0.004 & 
         tabla_completa_sex$delta_RMSEA[2] <= 0.050, "Yes", "No"),
  # Scalar: ΔCFI ≥ -0.004 Y ΔRMSEA ≤ 0.010
  ifelse(tabla_completa_sex$delta_CFI[3] >= -0.004 & 
         tabla_completa_sex$delta_RMSEA[3] <= 0.010, "Yes", "No")
)
```

## Final table formatting
```{r}
tabla_completa_cnt_export <- tabla_completa_cnt %>%
  select(Type, Model, chisq, df, CFI, TLI, RMSEA,
         delta_chisq, delta_df, 
         delta_CFI_display, delta_RMSEA_display,
         p_value_delta_chisq, Decision) %>%
  rename(delta_CFI = delta_CFI_display,
         delta_RMSEA = delta_RMSEA_display)

tabla_completa_sex_export <- tabla_completa_sex %>%
  select(Type, Model, chisq, df, CFI, TLI, RMSEA,
         delta_chisq, delta_df, 
         delta_CFI_display, delta_RMSEA_display,
         p_value_delta_chisq, Decision) %>%
  rename(delta_CFI = delta_CFI_display,
         delta_RMSEA = delta_RMSEA_display)

tabla_unificada_icils <- bind_rows(tabla_completa_cnt_export, tabla_completa_sex_export)

```

```{r}

colnames(tabla_unificada_icils) <- c(
  "Type",
  "Model", 
  "\u03C7\u00B2",      # χ²
  "df",
  "CFI",
  "TLI",
  "RMSEA",
  "\u0394\u03C7\u00B2", # Δχ²
  "\u0394df",          # Δdf
  "\u0394CFI",         # ΔCFI
  "\u0394RMSEA",       # ΔRMSEA
  "p",
  "Decision"
)

ft_icils <- flextable(tabla_unificada_icils) %>%
  colformat_double(j = c("\u03C7\u00B2", "df", "\u0394\u03C7\u00B2", "\u0394df"), digits = 0) %>%
  colformat_double(j = c("CFI", "TLI", "RMSEA", "\u0394CFI", "\u0394RMSEA", "p"), digits = 3) %>%  border_remove() %>%
  hline_top(border = officer::fp_border(width = 2), part = "all") %>%
  hline_bottom(border = officer::fp_border(width = 2), part = "all") %>%
  hline(i = 1, border = officer::fp_border(width = 1.5), part = "header") %>%
  bold(part = "header") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 9, part = "body") %>%
  fontsize(size = 10, part = "header") %>%
  padding(padding = 3, part = "all") %>%
  align(align = "center", part = "header") %>%
  align(align = "left", part = "body", j = 1:2) %>%
  align(align = "center", part = "body", j = 3:ncol(tabla_unificada_icils)) %>%
  autofit() %>%
  fit_to_width(max_width = 7.5)

# Which cols to color
delta_cfi_col <- which(colnames(tabla_unificada_icils) == "\u0394CFI")
delta_rmsea_col <- which(colnames(tabla_unificada_icils) == "\u0394RMSEA")
decision_col <- which(colnames(tabla_unificada_icils) == "Decision")
# Which rows to color
rows_cfi_fail <- which(!grepl("<", tabla_unificada_icils[[delta_cfi_col]]) & 
                       !is.na(tabla_unificada_icils[[delta_cfi_col]]))
rows_rmsea_fail <- which(!grepl("<", tabla_unificada_icils[[delta_rmsea_col]]) & 
                         !is.na(tabla_unificada_icils[[delta_rmsea_col]]))
rows_decision_no <- which(tabla_unificada_icils$Decision == "No")
# Red when dont achieve invariance
if(length(rows_cfi_fail) > 0) {
  ft_icils <- ft_icils %>%
    bg(i = rows_cfi_fail, j = delta_cfi_col, bg = "#FFCCCB") %>%
    color(i = rows_cfi_fail, j = delta_cfi_col, color = "#8B0000")
}

if(length(rows_rmsea_fail) > 0) {
  ft_icils <- ft_icils %>%
    bg(i = rows_rmsea_fail, j = delta_rmsea_col, bg = "#FFCCCB") %>%
    color(i = rows_rmsea_fail, j = delta_rmsea_col, color = "#8B0000")
}

if(length(rows_decision_no) > 0) {
  ft_icils <- ft_icils %>%
    bg(i = rows_decision_no, j = decision_col, bg = "#FFCCCB") %>%
    color(i = rows_decision_no, j = decision_col, color = "#8B0000") %>%
    bold(i = rows_decision_no, j = decision_col)
}

ft_icils
```

```{r}
#| eval: false

# Save
save_as_docx(ft_icils, path = "output/figures/tables/invariance_table_icils.docx")
save_as_image(ft_icils, path = "output/figures/tables/invariance_table_icils.png")
save_as_image(ft_icils, path = "output/article/images/invariance_table_icils.png")
saveRDS(ft_icils, "./output/article/tables/invariance_table_icils.rds")

```

# Descriptive analysis

## Latent means prediction

```{r}

fscores_df <- lavPredict(
  scalar_cnt,
  type = "lv", 
  method = "EBM",
  assemble = TRUE,
  append.data = TRUE
)

fscores_df |>
  summarise(
    total = n(),
    mean_gen_dse = mean(gen_dse, na.rm = TRUE),
    mean_spec_dse = mean(spec_dse, na.rm = TRUE)
  )

```

## Index and latent means correspondence analysis

### Index construction

```{r}

fscores_df <- fscores_df |> 
  mutate(
    gen_dse_index = rowMeans(pick(search_info, evaluate_info, install_app, 
                                  share_content, write_text, insert_image, edit_image,
                                  change_settings),
                             na.rm = TRUE),
                             
    spec_dse_index = rowMeans(pick(develop_webpage, programming, visual_coding), 
                              na.rm = TRUE)
  )
```

### Correlation index vs. latent mean

```{r}
cor_gen <- cor(fscores_df$gen_dse_index, fscores_df$gen_dse, 
               use = "complete.obs")
cor_spec <- cor(fscores_df$spec_dse_index, fscores_df$spec_dse, 
                use = "complete.obs")

cnt_label_map <- icils |> 
  select(CNTRY) |>
  mutate(label = to_label(CNTRY)) |> 
  distinct(CNTRY, .keep_all = TRUE)

n_global <- sum(!is.na(fscores_df$gen_dse))

global_row <- tibble(
  CNTRY = "GLOBAL",
  label = "Overall",
  N = n_global,
  `Gen_dse_correlations` = cor_gen,
  `Spec_dse_correlations` = cor_spec
)

cnt_cors <- fscores_df |>
  filter(!is.na(gen_dse)) |>
  group_by(CNTRY) |>
  summarise(
    N = n(),
    `Gen_dse_correlations` = cor(gen_dse_index, gen_dse, use = "complete.obs"),
    `Spec_dse_correlations` = cor(spec_dse_index, spec_dse, use = "complete.obs")
  ) |> 
  left_join(cnt_label_map, by = "CNTRY")

final_table <- bind_rows(global_row, cnt_cors) |> select(- CNTRY)

# knitr::kable(final_table, digits = 3, caption = "Index and latent means correlations")
```

### Exporting the table (for appendix)
```{r}

ft_cors <- flextable(final_table) |> 
  set_header_labels(
    label = "Country",
    N = "N",
    Gen_dse_correlations = "General DSE Correlation",
    Spec_dse_correlations = "Specialized DSE Correlation"
  ) |> 
  
  colformat_double(j = c("Gen_dse_correlations", "Spec_dse_correlations"), digits = 3) |> 
  
  border_remove() |> 
  hline_top(border = fp_border(width = 2), part = "all") |> 
  hline_bottom(border = fp_border(width = 2), part = "all") |> 
  hline(i = 1, border = fp_border(width = 1.5), part = "header") |> 
  
  bold(part = "header") |> 
  font(fontname = "Times New Roman", part = "all") |> 
  fontsize(size = 9, part = "body") |> 
  fontsize(size = 10, part = "header") |> 
  
  padding(padding = 3, part = "all") |> 
  align(align = "center", part = "header") |> 
  align(align = "left", part = "body", j = 1:2) |> 
  align(align = "center", part = "body", j = 3:4) |> 
  
  bold(i = 1, j = 1:4, part = "body") |> 
  autofit() |> 
  fit_to_width(max_width = 7.5)

ft_cors
```

```{r}
#| eval: false

saveRDS(ft_cors, "./output/article/tables/index_latent_means_icils.rds")
save_as_docx(ft_cors, path = "./output/figures/tables/index_latent_means_icils.docx")
save_as_image(ft_cors, path = "./output/figures/tables/index_latent_means_icils.png")
```

## Visualizations

### Index construction
```{r}
icils <- icils |> 
  mutate(
    gen_dse_index = rowMeans(pick(search_info, evaluate_info, install_app, 
                                  share_content, write_text, insert_image, edit_image,
                                  change_settings),
                             na.rm = TRUE),
                             
    spec_dse_index = rowMeans(pick(develop_webpage, programming, visual_coding), 
                              na.rm = TRUE)
  )
```

### Country comparison

```{r}

# Create data with country means
beeswarm_plot <- icils |>
  group_by(CNTRY)|>
  summarise(mean_geneff = weighted.mean(gen_dse_index, w = SENWT, na.rm = T),
            mean_speceff = weighted.mean(spec_dse_index, w = SENWT, na.rm = T))|>
  ungroup()|>
  pivot_longer(cols = 2:3,
               names_to = "type",
               values_to = "score")

# Sort and extract labels
beeswarm_labels <- 
  rbind(
  beeswarm_plot%>%
  group_by(type) %>%
  arrange(score) %>%
  slice(c(1:3, (n() - 2):n())) %>%
  ungroup() %>%
  mutate(CNTRY = to_label(CNTRY)),
  
  beeswarm_plot%>%
    filter(CNTRY %in% c(""))%>%
    mutate(CNTRY = to_label(CNTRY))%>%
    slice(-3)
  )

# Plot
icils_cnt_plot <- ggplot(beeswarm_plot, aes(y = score, x = type, color = type)) +
  geom_beeswarm(size = 2) +
  geom_text_repel(
    data = beeswarm_labels,
    aes(label = CNTRY),
    size = 3,
    color = "black",
    box.padding = 1.0,        
    point.padding = 0.1,      
    min.segment.length = 0,   
    max.overlaps = Inf,       
    force = 3,                
    force_pull = 0.3,         
    direction = "both",       
    seed = 999              
  ) +
  scale_color_manual(values = c("#5f5758","#fe3057")) +
  theme_minimal() +
  theme(
    legend.title = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    legend.position = "none",
    plot.caption = element_text(size = 15),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
  ) +
  scale_x_discrete(labels = c("General DSE", "Specialized DSE")) +
  scale_y_continuous(breaks = c(2,2.5,3,3.5,4), limits = c(2,4))

icils_cnt_plot
```

### Gender gap country comparison

```{r}
# Create data in long format with gender means
gender_plot <- icils |>  
  group_by(CNTRY, sex) |>  
  summarise(
    "General DSE" = weighted.mean(gen_dse_index, w = SENWT, na.rm = TRUE),
    "Specialized DSE" = weighted.mean(spec_dse_index, w = SENWT, na.rm = TRUE)
  ) |>  
  pivot_longer(
    cols = 3:4,
    names_to = "type",
    values_to = "score"
  ) |>  
  ungroup() |>  
  mutate(sex = to_label(sex)) |>  
  drop_na()

# Extract labels
country_labels_map <- gender_plot %>%
  distinct(CNTRY) %>%
  mutate(
    CNT_code = as.character(CNTRY),
    CNT_label = to_label(CNTRY)
  ) %>%
  select(CNT_code, CNT_label)

# Reorder the data by score for each type
gender_plot <- gender_plot %>%
  mutate(CNT_code = as.character(CNTRY)) %>%
  left_join(country_labels_map, by = "CNT_code") %>%
  group_by(type) %>%
  arrange(score) %>%
  mutate(CNT_label = factor(CNT_label, levels = unique(CNT_label))) %>%
  ungroup()


icils_gender_plot <- ggplot(gender_plot, aes(y = CNT_label, x = score)) + 
  facet_grid(~type, scales = "free_y") +
  geom_line(aes(group = CNT_label), size = 0.5) +  
  geom_point(aes(color = sex, shape = sex), size = 2) +  
  scale_color_manual(
    name = "Gender",
    labels = c("Female", "Male"),
    values = c("Female" = "#FE3057", "Male" = "#5f5758")) +
  scale_shape_manual(
    name = "Gender",
    labels = c("Female", "Male"),
    values = c("Female" = 16, "Male" = 17)) + 
  theme_minimal() +  
  theme(
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    legend.position = "top",
    axis.text.y = element_text(size = 8, color = "black"),
    plot.margin = margin(10, 10, 10, 10),
    panel.spacing = unit(1.5, "lines"),
    text = element_text(size = 12),
    strip.text = element_text(size = 12),
    legend.text = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
  ) +
  coord_cartesian(clip = "off") +
  scale_x_continuous(breaks = c(2, 2.5, 3, 3.5, 4), limits = c(1.8, 4))

icils_gender_plot
```

## Function for HD saving

```{r}
export_plot <- function(plot_object, 
                       filename,
                       format = c("png", "pdf", "tiff", "eps", "svg"),
                       width = 10,
                       height = 7,
                       dpi = 300,
                       units = "in",
                       bg = "white",
                       path = "./output/figures/") {
  
  format <- match.arg(format)
  
  full_filename <- paste0(filename, ".", format)
  full_path <- file.path(path, full_filename)
  

  ggsave(
    filename = full_path,
    plot = plot_object,
    width = width,
    height = height,
    dpi = dpi,
    units = units,
    bg = bg,
    device = format
  )
  
  message(paste0("✓ Gráfico exportado: ", full_path))
}
```

## Function for multiformat HD saving

```{r}
export_plot_multi <- function(plot_object,
                              filename,
                              formats = c("png", "pdf", "tiff"),
                              width = 10,
                              height = 7,
                              dpi = 300,
                              units = "in",
                              bg = "white",
                              path = "./output/figures/") {
  
  purrr::walk(formats, ~export_plot(
    plot_object = plot_object,
    filename = filename,
    format = .x,
    width = width,
    height = height,
    dpi = dpi,
    units = units,
    bg = bg,
    path = path
  ))
}
```


## Saving the plots in multiple formats

```{r}
#| eval: false

export_plot_multi(icils_cnt_plot, "icils_country_dse_comparison", 
                  formats = c("png", "pdf", "tiff"), width = 10, height = 6, dpi = 600, path = "./output/figures/plots/")

export_plot_multi(icils_gender_plot, "icils_gender_dse_gaps", 
                  formats = c("png", "pdf", "tiff"), width = 10, height = 6, dpi = 600, path = "./output/figures/plots/")
# Article
export_plot_multi(icils_cnt_plot, "icils_country_dse_comparison", 
                  formats = c("png"), width = 10, height = 6, dpi = 600, path = "./output/article/images/")

export_plot_multi(icils_gender_plot, "icils_gender_dse_gaps", 
                  formats = c("png"), width = 10, height = 6, dpi = 600, path = "./output/article/images/")

```

# Loadings and chi analysis (appendix)

```{r}

all_params <- parameterEstimates(scalar_cnt, standardized = TRUE)

factor_loadings_country <- all_params %>%
  filter(op == "=~")

test_stats <- lavInspect(scalar_cnt, "test")
chisq_values <- test_stats$scaled.shifted$stat.group
country_names_lavaan <- attr(test_stats, "info")$group.label

chisq_map <- data.frame(
  CNTRY = country_names_lavaan,
  Chisqr = as.numeric(chisq_values)
)

label_dictionary <- icils %>%
  select(CNTRY) %>%
  distinct() %>%
  mutate(
    label_text = as.character(haven::as_factor(CNTRY)) 
  )

n_countries <- icils |> 
  group_by(CNTRY) |> 
  select(CNTRY) |> 
  mutate(N = n()) |> 
  distinct(CNTRY, .keep_all = TRUE)

groups_code <- lavInspect(scalar_cnt, "group.label")

matched_labels <- label_dictionary$label_text[match(groups_code, label_dictionary$CNTRY)]

country_map <- data.frame(
  group = 1:length(groups_code),
  CNTRY = groups_code,
  Country = matched_labels
)

factor_loadings_final <- factor_loadings_country %>%
  left_join(country_map, by = "group") %>%
  left_join(n_countries, by = "CNTRY") %>%
  left_join(chisq_map, by = "CNTRY") %>%
  select(Country, Item = rhs, Factor = lhs, Loading = std.all, N, Chisqr)


```

## Table for appendix

```{r}
loadings_table <- factor_loadings_final %>%
  select(Country, Factor, Item, Loading, N, Chisqr) %>%
  arrange(desc(Chisqr), Factor)

borde_grueso <- fp_border(width = 2)
borde_medio  <- fp_border(width = 1.5)
borde_gris   <- fp_border(color = "grey80", width = 0.5)

ft_loadings <- flextable(loadings_table) %>%
  
  set_header_labels(
    Country = "Country",
    Factor  = "Latent Factor",
    Item    = "Item",
    Loading = "Std. Loading",
    N       = "N",
    Chisqr  = "\u03C7\u00B2"
  ) %>%
  
  merge_v(j = c("Country", "Factor", "N", "Chisqr")) %>%
  
  colformat_double(j = "Loading", digits = 3) %>%
  colformat_double(j = "Chisqr", digits = 2) %>%

  colformat_double(j = "N", digits = 0) %>% 
  
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 9, part = "body") %>%
  fontsize(size = 10, part = "header") %>%
  bold(part = "header") %>%
  
  padding(padding = 3, part = "all") %>%
  align(align = "center", part = "header") %>%
  
  align(align = "center", j = c("Factor", "Loading", "N", "Chisqr"), part = "body") %>% 
  align(align = "left", j = c("Country", "Item"), part = "body") %>%
  
  valign(valign = "center", j = c("Country", "Factor", "N", "Chisqr"), part = "body") %>%

  border_remove() %>%
  
  hline(i = ~ !duplicated(Country, fromLast = TRUE), 
        border = borde_gris) %>%
  
  hline_top(border = borde_grueso, part = "all") %>%
  hline_bottom(border = borde_grueso, part = "all") %>%
  hline(part = "header", border = borde_medio) %>%
  
  autofit() %>%
  fit_to_width(max_width = 7.5)

ft_loadings
```

### Table for report

```{r}

datatable(
  factor_loadings_final, 
  rownames = FALSE,
  filter = 'top',
  options = list(
    pageLength = 15,
    autoWidth = TRUE
  ),
  caption = 'Interactive table: Loadings and Chi square by country'
) %>%
  formatRound(columns = 'Loading', digits = 3) %>% 
  formatRound(columns = 'Chisqr', digits = 2)

```

```{r}
#| eval: false

# Saving table
saveRDS(ft_loadings, "./output/article/tables/loadings_chi_icils.rds")
save_as_docx(ft_loadings, path = "./output/figures/tables/loadings_chi_icils.docx")
save_as_image(ft_loadings, path = "./output/figures/tables/loadings_chi_icils.png")
```