---
title: "PISA analysis script"
author: "Ismael Aguayo"
date: today
lang: en
fontsize: 14pt
format:
  html:
    code-fold: false
    toc: true
    toc-location: left
    toc-depth: 2
    toc-expand: 1
    toc-title: Contents
    number-sections: true
    number-depth: 3
    theme:
      - united
    code-link: true
    title-block-banner: true
editor_options: 
  chunk_output_type: console
---

# Enviroment setting

```{r}
#install.packages("pacman")
library(pacman)

pacman::p_load(
  dplyr, haven, sjlabelled,  psych,  purrr,  tidyr,  sjPlot,  ggplot2, 
  parameters,  table1,  car,  beeswarm,  lme4, scales, ggrepel, corrplot,
  ggtext, patchwork, lavaan, semTools, DT, knitr, kableExtra, flextable, officer)

options(scipen = 999)
rm(list = ls()) 

pisa <- readRDS(here::here("input", "data", "proc_data", "pisa_comparable.rds"))

```

# Pooled model

```{r}
model_syntax <- '
  gen_dse =~ search_info + evaluate_info + collect_data + 
  explain_content + share_content + pair_collab + 
  write_text + create_media + change_settings
  spec_dse =~ programming + develop_webpage + identify_error + identify_app + logical_solution
  '

m_pooled <- cfa(model = model_syntax, 
              data = pisa,
              estimator = "WLSMV",
              missing = 'pairwise',
              sampling.weights = "SENWT",
              ordered = T)

```

### Fit

```{r}
fit_pisa <- fitMeasures(m_pooled,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

knitr::kable(fit_pisa, digits = 3)
```

Given the poor fit of the initial model (RMSEA of .12), we look at the modindices and the residual covariance matrix.

### Modification indices and residual covariances

```{r}
mod_indices <- modindices(m_pooled, sort = TRUE, maximum.number = 10)

knitr::kable(mod_indices, digits = 3)

matriz_residuales <- resid(m_pooled, type = "standardized")$cov

corrplot(matriz_residuales, 
         is.corr = FALSE,       
         method = "color", 
         tl.col = "black",      
         cl.pos = "r",          
         type = "lower")
```

First, we detect  a couple of problematic items. Identify app charge to the general factor with a mi of 76,502 (very high), and shows strong residual covariance with a lot of items (change settings, identify error, logical solution, programming). We decide to remove that item, because is causing too much noise in the model and possibly dont adjust with the programing-related tasks of the specialized DSE factor.

### Re-especification

```{r}
model_syntax_2 <- '
  gen_dse =~ search_info + evaluate_info + collect_data + 
  explain_content + share_content + pair_collab + 
  write_text + create_media + change_settings
  spec_dse =~ programming + develop_webpage + identify_error + logical_solution
  '

m_pooled_2 <- cfa(model = model_syntax_2, 
              data = pisa,
              estimator = "WLSMV",
              missing = 'pairwise',
              sampling.weights = "SENWT",
              ordered = T)
```

### Fit

```{r}
fit_pisa <- fitMeasures(m_pooled_2,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

knitr::kable(fit_pisa, digits = 3)
```

The fit of the model improved considerably (RMSEA = .094), but still shows space for adjustment. Therefore, we look again to the modindices and the residual standarized covariance matrix.

### Modification indices and residual covariances

```{r}
mod_indices <- modindices(m_pooled_2, sort = TRUE, maximum.number = 10)

knitr::kable(mod_indices, digits = 3)

matriz_residuales <- resid(m_pooled_2, type = "standardized")$cov

corrplot(matriz_residuales, 
         is.corr = FALSE,       
         method = "color", 
         tl.col = "black",      
         cl.pos = "r",          
         type = "lower")
```

The item of develop webpage self-efficacy shows a big cross-loading to the general factor (mi=39,014). Looking at the residuals, it has strong residual covariance with collect data, create media and change settings, and a negative one with items from the specialized factor (i.e. logical solution and identify error). This item will be also be removed. Summing up to the technical reasons, the growing democratization of web development impulsed by the user-friendly platforms might have changed the perception of this item by the students in this cycle. In comparison with ICILS, this item refers to a webpage or a blog, it could be that students relate blogs with more non-technical platforms like fandom.

### Re-especification 2

```{r}
model_syntax_3 <- '
  gen_dse =~ search_info + evaluate_info + collect_data + 
  explain_content + share_content + pair_collab + 
  write_text + create_media + change_settings
  spec_dse =~ programming + identify_error + logical_solution
  '

m_pooled_3 <- cfa(model = model_syntax_3, 
              data = pisa,
              estimator = "WLSMV",
              missing = 'pairwise',
              sampling.weights = "SENWT",
              ordered = T)
```

### Fit

```{r}
fit_pisa <- fitMeasures(m_pooled_3,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

knitr::kable(fit_pisa, digits = 3)
```

With this two adjustments, the model is good to go with an RMSEA of .068. Given this, we continue with further analysis.

```{r}
# Releasing a bit of ram
rm(m_pooled, m_pooled_2)
```

### Item information

```{r}

std_params <- standardizedsolution(m_pooled_3)


items_table <- std_params %>% 
  filter(op == "=~") %>% 
  select(Factor = lhs, 
         Indicator = rhs, 
         Loading = est.std, 
         SE = se, 
         p_value = pvalue) %>% 
  mutate(R2_Explained = Loading^2)

knitr::kable(items_table, digits = 3, caption = "Loadings and R squared")

```

### Reliability analysis

```{r}
omega_cat <- compRelSEM(m_pooled_3)

ave_values <- AVE(m_pooled_3)

reliability_table <- data.frame(
  Factor = names(omega_cat),
  Omega_Categorico = as.numeric(omega_cat),
  AVE = as.numeric(ave_values)
)

knitr::kable(reliability_table, digits = 3, caption = "Fiabilidad (Omega Categórico) y Validez (AVE)")
```

### Inter-factor correlation

```{r}

std_params %>% 
  filter(op == "~~",      
         lhs != rhs) %>% 
  select(Factor1 = lhs, Factor2 = rhs, Correlation = est.std, 
         SE = se, Z = z, `p-value` = pvalue)

```

# Invariance analysis

## Country invariance

```{r}
#| warning: false

config_cnt <- cfa(model = model_syntax_3, 
              data = pisa,
              estimator = "WLSMV",
              group = "CNT",
              sampling.weights = "SENWT",
              sampling.weights.normalization = "group",
              missing = 'pairwise',
              ordered = T)

metric_cnt <- cfa(model = model_syntax_3, 
              data = pisa,
              estimator = "WLSMV",
              group = "CNT",
              sampling.weights = "SENWT",
              sampling.weights.normalization = "group",
              ordered = T,
              missing = 'pairwise',
              group.equal= "loadings")

scalar_cnt <- cfa(model = model_syntax_3, 
              data = pisa,
              estimator = "WLSMV",
              group = "CNT",
              sampling.weights = "SENWT",
              sampling.weights.normalization = "group",
              ordered = T,
              missing = 'pairwise',
              group.equal= c("loadings", "intercepts"))

```

### Fit measures estracting

```{r}
fitmeasures_config_cnt <- fitMeasures(config_cnt,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

fitmeasures_metric_cnt <- fitMeasures(metric_cnt,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

fitmeasures_scalar_cnt <- fitMeasures(scalar_cnt,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))
```

### Table processing

```{r}
# Dataframe consolidado

tabla_completa_cnt <- data.frame(
  Type = c("Country", "", ""),
  Model = c("1. Configural", "2. Metric", "3. Scalar"),
  chisq = c(fitmeasures_config_cnt["chisq"], fitmeasures_metric_cnt["chisq"], fitmeasures_scalar_cnt["chisq"]),
  df = c(fitmeasures_config_cnt["df"], fitmeasures_metric_cnt["df"], fitmeasures_scalar_cnt["df"]),
  CFI = c(fitmeasures_config_cnt["cfi"], fitmeasures_metric_cnt["cfi"], fitmeasures_scalar_cnt["cfi"]),
  TLI = c(fitmeasures_config_cnt["tli"], fitmeasures_metric_cnt["tli"], fitmeasures_scalar_cnt["tli"]),
  RMSEA = c(fitmeasures_config_cnt["rmsea"], fitmeasures_metric_cnt["rmsea"], fitmeasures_scalar_cnt["rmsea"])
)

# Cálculo deltas

tabla_completa_cnt$delta_chisq <- c(NA, diff(tabla_completa_cnt$chisq))
tabla_completa_cnt$delta_df    <- c(NA, diff(tabla_completa_cnt$df))
tabla_completa_cnt$delta_CFI   <- c(NA, diff(tabla_completa_cnt$CFI))
tabla_completa_cnt$delta_RMSEA <- c(NA, diff(tabla_completa_cnt$RMSEA))

# Display para los deltas 

tabla_completa_cnt$delta_CFI_display <- sapply(1:nrow(tabla_completa_cnt), function(i) {
  if (i == 1) return(NA)
  
  if (tabla_completa_cnt$delta_CFI[i] >= -0.004) {
    return("< -0.004")
  } else {
    return(sprintf("%.3f", tabla_completa_cnt$delta_CFI[i]))
  }
})

tabla_completa_cnt$delta_RMSEA_display <- sapply(1:nrow(tabla_completa_cnt), function(i) {
  if (i == 1) return(NA)
  
  umbral <- if (i == 2) 0.050 else 0.010
  
  if (tabla_completa_cnt$delta_RMSEA[i] <= umbral) {
    return(paste0("< ", umbral))
  } else {
    return(sprintf("%.3f", tabla_completa_cnt$delta_RMSEA[i]))
  }
})

# ANOVA para significancia

anova_cnt <- anova(config_cnt, metric_cnt, scalar_cnt)

tabla_completa_cnt$p_value_delta_chisq <- case_when(anova_cnt$`Pr(>Chisq)`< 0.01 ~ "< 0.01",
                                                  anova_cnt$`Pr(>Chisq)`< 0.05 ~ "< 0.05",
                                                  anova_cnt$`Pr(>Chisq)`> 0.05 ~ sprintf("%.2f", anova_cnt$`Pr(>Chisq)`))
tabla_completa_cnt$Decision <- c(
  NA,  # Configural no tiene comparación previa
  # Metric: ΔCFI ≥ -0.004 Y ΔRMSEA ≤ 0.050
  ifelse(tabla_completa_cnt$delta_CFI[2] >= -0.004 & 
         tabla_completa_cnt$delta_RMSEA[2] <= 0.050, "Yes", "No"),
  # Scalar: ΔCFI ≥ -0.004 Y ΔRMSEA ≤ 0.010
  ifelse(tabla_completa_cnt$delta_CFI[3] >= -0.004 & 
         tabla_completa_cnt$delta_RMSEA[3] <= 0.010, "Yes", "No")
)
```

## Gender invariance

### Model fitting
```{r}
#| warning: false

config_sex <- cfa(model = model_syntax_3, 
              data = pisa,
              estimator = "WLSMV",
              group = "sex",
              sampling.weights = "SENWT",
              sampling.weights.normalization = "group",
              missing = 'pairwise',
              ordered = T)

metric_sex <- cfa(model = model_syntax_3, 
              data = pisa,
              estimator = "WLSMV",
              group = "sex",
              sampling.weights = "SENWT",
              sampling.weights.normalization = "group",
              ordered = T,
              missing = 'pairwise',
              group.equal= "loadings")

scalar_sex <- cfa(model = model_syntax_3, 
              data = pisa,
              estimator = "WLSMV",
              group = "sex",
              sampling.weights = "SENWT",
              sampling.weights.normalization = "group",
              ordered = T,
              missing = 'pairwise',
              group.equal= c("loadings", "intercepts"))

```

### Fit measures stracting

```{r}
fitmeasures_config_sex <- fitMeasures(config_sex,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

fitmeasures_metric_sex <- fitMeasures(metric_sex,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

fitmeasures_scalar_sex <- fitMeasures(scalar_sex,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))
```

### Table processing
```{r}
# Dataframe consolidado

tabla_completa_sex <- data.frame(
  Type = c("Gender", "", ""),
  Model = c("1. Configural", "2. Metric", "3. Scalar"),
  chisq = c(fitmeasures_config_sex["chisq"], fitmeasures_metric_sex["chisq"], fitmeasures_scalar_sex["chisq"]),
  df = c(fitmeasures_config_sex["df"], fitmeasures_metric_sex["df"], fitmeasures_scalar_sex["df"]),
  CFI = c(fitmeasures_config_sex["cfi"], fitmeasures_metric_sex["cfi"], fitmeasures_scalar_sex["cfi"]),
  TLI = c(fitmeasures_config_sex["tli"], fitmeasures_metric_sex["tli"], fitmeasures_scalar_sex["tli"]),
  RMSEA = c(fitmeasures_config_sex["rmsea"], fitmeasures_metric_sex["rmsea"], fitmeasures_scalar_sex["rmsea"])
)

# Cálculo deltas

tabla_completa_sex$delta_chisq <- c(NA, diff(tabla_completa_sex$chisq))
tabla_completa_sex$delta_df    <- c(NA, diff(tabla_completa_sex$df))
tabla_completa_sex$delta_CFI   <- c(NA, diff(tabla_completa_sex$CFI))
tabla_completa_sex$delta_RMSEA <- c(NA, diff(tabla_completa_sex$RMSEA))

# Display para los deltas 

tabla_completa_sex$delta_CFI_display <- sapply(1:nrow(tabla_completa_sex), function(i) {
  if (i == 1) return(NA)
  
  if (tabla_completa_sex$delta_CFI[i] >= -0.004) {
    return("< -0.004")
  } else {
    return(sprintf("%.3f", tabla_completa_sex$delta_CFI[i]))
  }
})

tabla_completa_sex$delta_RMSEA_display <- sapply(1:nrow(tabla_completa_sex), function(i) {
  if (i == 1) return(NA)
  
  umbral <- if (i == 2) 0.050 else 0.010
  
  if (tabla_completa_sex$delta_RMSEA[i] <= umbral) {
    return(paste0("< ", umbral))
  } else {
    return(sprintf("%.3f", tabla_completa_sex$delta_RMSEA[i]))
  }
})

# ANOVA para significancia

anova_sex <- anova(config_sex, metric_sex, scalar_sex)

tabla_completa_sex$p_value_delta_chisq <- case_when(anova_sex$`Pr(>Chisq)`< 0.01 ~ "< 0.01",
                                                  anova_sex$`Pr(>Chisq)`< 0.05 ~ "< 0.05",
                                                  anova_sex$`Pr(>Chisq)`> 0.05 ~ sprintf("%.2f", anova_sex$`Pr(>Chisq)`))
tabla_completa_sex$Decision <- c(
  NA,  # Configural no tiene comparación previa
  # Métrico: ΔCFI ≥ -0.004 Y ΔRMSEA ≤ 0.050
  ifelse(tabla_completa_sex$delta_CFI[2] >= -0.004 & 
         tabla_completa_sex$delta_RMSEA[2] <= 0.050, "Yes", "No"),
  # Escalar: ΔCFI ≥ -0.004 Y ΔRMSEA ≤ 0.010
  ifelse(tabla_completa_sex$delta_CFI[3] >= -0.004 & 
         tabla_completa_sex$delta_RMSEA[3] <= 0.010, "Yes", "No")
)
```

## Final table formatting
```{r}
tabla_completa_cnt_export <- tabla_completa_cnt %>%
  select(Type, Model, chisq, df, CFI, TLI, RMSEA,
         delta_chisq, delta_df, 
         delta_CFI_display, delta_RMSEA_display,
         p_value_delta_chisq, Decision) %>%
  rename(delta_CFI = delta_CFI_display,
         delta_RMSEA = delta_RMSEA_display)

tabla_completa_sex_export <- tabla_completa_sex %>%
  select(Type, Model, chisq, df, CFI, TLI, RMSEA,
         delta_chisq, delta_df, 
         delta_CFI_display, delta_RMSEA_display,
         p_value_delta_chisq, Decision) %>%
  rename(delta_CFI = delta_CFI_display,
         delta_RMSEA = delta_RMSEA_display)

tabla_unificada_pisa <- bind_rows(tabla_completa_cnt_export, tabla_completa_sex_export)

```

```{r}

colnames(tabla_unificada_pisa) <- c(
  "Type",
  "Model", 
  "\u03C7\u00B2",      # χ²
  "df",
  "CFI",
  "TLI",
  "RMSEA",
  "\u0394\u03C7\u00B2", # Δχ²
  "\u0394df",          # Δdf
  "\u0394CFI",         # ΔCFI
  "\u0394RMSEA",       # ΔRMSEA
  "p",
  "Decision"
)

ft_pisa <- flextable(tabla_unificada_pisa) %>%
# Formatear números: 0 decimales para χ², df, Δχ², Δdf
  colformat_double(j = c("\u03C7\u00B2", "df", "\u0394\u03C7\u00B2", "\u0394df"), digits = 0) %>%
  # 3 decimales para el resto de columnas numéricas
  colformat_double(j = c("CFI", "TLI", "RMSEA", "\u0394CFI", "\u0394RMSEA", "p"), digits = 3) %>%  border_remove() %>%
  hline_top(border = officer::fp_border(width = 2), part = "all") %>%
  hline_bottom(border = officer::fp_border(width = 2), part = "all") %>%
  hline(i = 1, border = officer::fp_border(width = 1.5), part = "header") %>%
  bold(part = "header") %>%
  font(fontname = "Calibri", part = "all") %>%
  fontsize(size = 9, part = "body") %>%
  fontsize(size = 10, part = "header") %>%
  padding(padding = 3, part = "all") %>%
  align(align = "center", part = "header") %>%
  align(align = "left", part = "body", j = 1:2) %>%
  align(align = "center", part = "body", j = 3:ncol(tabla_unificada_pisa)) %>%
  autofit() %>%
  fit_to_width(max_width = 7.5)

# --- LOGICA DE COLOREO AUTOMÁTICA --- #

# Detectamos la columna a colorear
delta_cfi_col <- which(colnames(tabla_unificada_pisa) == "\u0394CFI")
delta_rmsea_col <- which(colnames(tabla_unificada_pisa) == "\u0394RMSEA")
decision_col <- which(colnames(tabla_unificada_pisa) == "Decision")
# Detectamos la fila a colorear (las que no tienen el operador menor que)
rows_cfi_fail <- which(!grepl("<", tabla_unificada_pisa[[delta_cfi_col]]) & 
                       !is.na(tabla_unificada_pisa[[delta_cfi_col]]))
rows_rmsea_fail <- which(!grepl("<", tabla_unificada_pisa[[delta_rmsea_col]]) & 
                         !is.na(tabla_unificada_pisa[[delta_rmsea_col]]))
rows_decision_no <- which(tabla_unificada_pisa$Decision == "No")
# Coloreamos según corresponda
if(length(rows_cfi_fail) > 0) {
  ft_pisa <- ft_pisa %>%
    bg(i = rows_cfi_fail, j = delta_cfi_col, bg = "#FFCCCB") %>%
    color(i = rows_cfi_fail, j = delta_cfi_col, color = "#8B0000")
}

if(length(rows_rmsea_fail) > 0) {
  ft_pisa <- ft_pisa %>%
    bg(i = rows_rmsea_fail, j = delta_rmsea_col, bg = "#FFCCCB") %>%
    color(i = rows_rmsea_fail, j = delta_rmsea_col, color = "#8B0000")
}

if(length(rows_decision_no) > 0) {
  ft_pisa <- ft_pisa %>%
    bg(i = rows_decision_no, j = decision_col, bg = "#FFCCCB") %>%
    color(i = rows_decision_no, j = decision_col, color = "#8B0000") %>%
    bold(i = rows_decision_no, j = decision_col)
}

ft_pisa
```

```{r}
#| eval: false

# Guardar como docx y png
save_as_docx(ft_pisa, path = "output/figures/tables/invariance_table_pisa.docx")
save_as_image(ft_pisa, path = "output/figures/tables/invariance_table_pisa.png")
save_as_image(ft_pisa, path = "output/article/images/invariance_table_pisa.png")

```

# Descriptive analysis

```{r}

fscores_p <- lavPredict(fitgroup_p)

pisa$gen_dse_score <- NA_real_
pisa$spec_dse_score <- NA_real_

for (g in seq_along(fscores_p)) {
  # idx[[g]] contiene los índices de filas que lavaan usó para este grupo
  # fscores[[g]] contiene los factor scores para esas filas
  
  pisa[idx_p[[g]], "gen_dse_score"] <- fscores_p[[g]][, "gen_dse"]
  pisa[idx_p[[g]], "spec_dse_score"] <- fscores_p[[g]][, "spec_dse"]
}



```

Probar e investigar parámetros lavpredict newdata, method, transform, append.data y assemble. Luego, realizar visualizaciones para PISA con el modelo escalar con los índices si nada de esto funciona y si funciona con las medias latentes.