---
title: "PISA analysis script"
author: "Ismael Aguayo"
date: today
lang: en
fontsize: 14pt
format:
  html:
    code-fold: false
    toc: true
    toc-location: left
    toc-depth: 2
    toc-expand: 1
    toc-title: Contents
    number-sections: true
    number-depth: 3
    theme:
      - united
    code-link: true
    title-block-banner: true
editor_options: 
  chunk_output_type: console
---

# Enviroment setting

```{r}
#install.packages("pacman")
library(pacman)

pacman::p_load(
  dplyr, haven, sjlabelled,  psych,  purrr,  tidyr,  sjPlot,  ggplot2, 
  parameters,  table1,  car,  beeswarm,  lme4, scales, ggrepel, corrplot,
  ggtext, patchwork, lavaan, semTools, DT, knitr, kableExtra, flextable, officer, ggbeeswarm)

options(scipen = 999)
rm(list = ls()) 

pisa <- readRDS(here::here("input", "data", "proc_data", "pisa_comparable.rds"))

```

# Pooled model

```{r}
#| warning: false

model_syntax <- '
  gen_dse =~ search_info + evaluate_info + collect_data + 
  explain_content + share_content + pair_collab + 
  write_text + create_media + change_settings
  spec_dse =~ programming + develop_webpage + identify_error + identify_app + logical_solution
  '

m_pooled <- cfa(model = model_syntax, 
              data = pisa,
              estimator = "WLSMV",
              missing = 'pairwise',
              sampling.weights = "SENWT",
              ordered = T)

```

### Fit

```{r}
fit_pisa <- fitMeasures(m_pooled,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

knitr::kable(fit_pisa, digits = 3)
```

Given the poor fit of the initial model (RMSEA of .12), we look at the modindices and the residual covariance matrix.

### Modification indices and residual covariances

```{r}
mod_indices <- modindices(m_pooled, sort = TRUE, maximum.number = 10)

knitr::kable(mod_indices, digits = 3)

matriz_residuales <- resid(m_pooled, type = "standardized")$cov

corrplot(matriz_residuales, 
         is.corr = FALSE,       
         method = "color", 
         tl.col = "black",      
         cl.pos = "r",          
         type = "lower")
```

First, we detect  a couple of problematic items. Identify app charge to the general factor with a mi of 76,502 (very high), and shows strong residual covariance with a lot of items (change settings, identify error, logical solution, programming). We decide to remove that item, because is causing too much noise in the model and possibly dont adjust with the programing-related tasks of the specialized DSE factor.

### Re-especification

```{r}
#| warning: false

model_syntax_2 <- '
  gen_dse =~ search_info + evaluate_info + collect_data + 
  explain_content + share_content + pair_collab + 
  write_text + create_media + change_settings
  spec_dse =~ programming + develop_webpage + identify_error + logical_solution
  '

m_pooled_2 <- cfa(model = model_syntax_2, 
              data = pisa,
              estimator = "WLSMV",
              missing = 'pairwise',
              sampling.weights = "SENWT",
              ordered = T)
```

### Fit

```{r}
fit_pisa <- fitMeasures(m_pooled_2,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

knitr::kable(fit_pisa, digits = 3)
```

The fit of the model improved considerably (RMSEA = .094), but still shows space for adjustment. Therefore, we look again to the modindices and the residual standarized covariance matrix.

### Modification indices and residual covariances

```{r}
mod_indices <- modindices(m_pooled_2, sort = TRUE, maximum.number = 10)

knitr::kable(mod_indices, digits = 3)

matriz_residuales <- resid(m_pooled_2, type = "standardized")$cov

corrplot(matriz_residuales, 
         is.corr = FALSE,       
         method = "color", 
         tl.col = "black",      
         cl.pos = "r",          
         type = "lower")
```

The item of develop webpage self-efficacy shows a big cross-loading to the general factor (mi=39,014). Looking at the residuals, it has strong residual covariance with collect data, create media and change settings, and a negative one with items from the specialized factor (i.e. logical solution and identify error). This item will be also be removed. Summing up to the technical reasons, the growing democratization of web development impulsed by the user-friendly platforms might have changed the perception of this item by the students in this cycle. In comparison with ICILS, this item refers to a webpage or a blog, it could be that students relate blogs with more non-technical platforms like fandom.

### Re-especification 2

```{r}
#| warning: false

model_syntax_3 <- '
  gen_dse =~ search_info + evaluate_info + collect_data + 
  explain_content + share_content + pair_collab + 
  write_text + create_media + change_settings
  spec_dse =~ programming + identify_error + logical_solution
  '

m_pooled_3 <- cfa(model = model_syntax_3, 
              data = pisa,
              estimator = "WLSMV",
              missing = 'pairwise',
              sampling.weights = "SENWT",
              ordered = T)
```

### Fit

```{r}
fit_pisa <- fitMeasures(m_pooled_3,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

knitr::kable(fit_pisa, digits = 3)
```

With this two adjustments, the model is good to go with an RMSEA of .068. Given this, we continue with further analysis.

```{r}
# Releasing a bit of ram
rm(m_pooled, m_pooled_2)
```

### Item information

```{r}

std_params <- standardizedsolution(m_pooled_3)


items_table <- std_params %>% 
  filter(op == "=~") %>% 
  select(Factor = lhs, 
         Indicator = rhs, 
         Loading = est.std, 
         SE = se, 
         p_value = pvalue) %>% 
  mutate(R2_Explained = Loading^2)

knitr::kable(items_table, digits = 3, caption = "Loadings and R squared")

```

### Reliability analysis

```{r}
omega_cat <- compRelSEM(m_pooled_3)

ave_values <- AVE(m_pooled_3)

reliability_table <- data.frame(
  Factor = names(omega_cat),
  Omega_Categorico = as.numeric(omega_cat),
  AVE = as.numeric(ave_values)
)

knitr::kable(reliability_table, digits = 3, caption = "Fiabilidad (Omega Categórico) y Validez (AVE)")
```

### Inter-factor correlation

```{r}

std_params %>% 
  filter(op == "~~",      
         lhs != rhs) %>% 
  select(Factor1 = lhs, Factor2 = rhs, Correlation = est.std, 
         SE = se, Z = z, `p-value` = pvalue)

```

# Invariance analysis

## Country invariance

```{r}
#| warning: false

config_cnt <- cfa(model = model_syntax_3, 
              data = pisa,
              estimator = "WLSMV",
              group = "CNT",
              sampling.weights = "SENWT",
              sampling.weights.normalization = "group",
              missing = 'pairwise',
              ordered = T)

metric_cnt <- cfa(model = model_syntax_3, 
              data = pisa,
              estimator = "WLSMV",
              group = "CNT",
              sampling.weights = "SENWT",
              sampling.weights.normalization = "group",
              ordered = T,
              missing = 'pairwise',
              group.equal= "loadings")

scalar_cnt <- cfa(model = model_syntax_3, 
              data = pisa,
              estimator = "WLSMV",
              group = "CNT",
              sampling.weights = "SENWT",
              sampling.weights.normalization = "group",
              ordered = T,
              missing = 'pairwise',
              group.equal= c("loadings", "intercepts"))

```

### Fit measures estracting

```{r}
fitmeasures_config_cnt <- fitMeasures(config_cnt,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

fitmeasures_metric_cnt <- fitMeasures(metric_cnt,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

fitmeasures_scalar_cnt <- fitMeasures(scalar_cnt,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))
```

### Table processing

```{r}

tabla_completa_cnt <- data.frame(
  Type = c("Country", "", ""),
  Model = c("1. Configural", "2. Metric", "3. Scalar"),
  chisq = c(fitmeasures_config_cnt["chisq"], fitmeasures_metric_cnt["chisq"], fitmeasures_scalar_cnt["chisq"]),
  df = c(fitmeasures_config_cnt["df"], fitmeasures_metric_cnt["df"], fitmeasures_scalar_cnt["df"]),
  CFI = c(fitmeasures_config_cnt["cfi"], fitmeasures_metric_cnt["cfi"], fitmeasures_scalar_cnt["cfi"]),
  TLI = c(fitmeasures_config_cnt["tli"], fitmeasures_metric_cnt["tli"], fitmeasures_scalar_cnt["tli"]),
  RMSEA = c(fitmeasures_config_cnt["rmsea"], fitmeasures_metric_cnt["rmsea"], fitmeasures_scalar_cnt["rmsea"])
)

# Deltas

tabla_completa_cnt$delta_chisq <- c(NA, diff(tabla_completa_cnt$chisq))
tabla_completa_cnt$delta_df    <- c(NA, diff(tabla_completa_cnt$df))
tabla_completa_cnt$delta_CFI   <- c(NA, diff(tabla_completa_cnt$CFI))
tabla_completa_cnt$delta_RMSEA <- c(NA, diff(tabla_completa_cnt$RMSEA))

# Deltas display

tabla_completa_cnt$delta_CFI_display <- sapply(1:nrow(tabla_completa_cnt), function(i) {
  if (i == 1) return(NA)
  
  if (tabla_completa_cnt$delta_CFI[i] >= -0.004) {
    return("< -0.004")
  } else {
    return(sprintf("%.3f", tabla_completa_cnt$delta_CFI[i]))
  }
})

tabla_completa_cnt$delta_RMSEA_display <- sapply(1:nrow(tabla_completa_cnt), function(i) {
  if (i == 1) return(NA)
  
  umbral <- if (i == 2) 0.050 else 0.010
  
  if (tabla_completa_cnt$delta_RMSEA[i] <= umbral) {
    return(paste0("< ", umbral))
  } else {
    return(sprintf("%.3f", tabla_completa_cnt$delta_RMSEA[i]))
  }
})

# ANOVA for significance

anova_cnt <- anova(config_cnt, metric_cnt, scalar_cnt)

tabla_completa_cnt$p_value_delta_chisq <- case_when(anova_cnt$`Pr(>Chisq)`< 0.01 ~ "< 0.01",
                                                  anova_cnt$`Pr(>Chisq)`< 0.05 ~ "< 0.05",
                                                  anova_cnt$`Pr(>Chisq)`> 0.05 ~ sprintf("%.2f", anova_cnt$`Pr(>Chisq)`))
tabla_completa_cnt$Decision <- c(
  NA,  # Configural
  # Metric: ΔCFI ≥ -0.004 Y ΔRMSEA ≤ 0.050
  ifelse(tabla_completa_cnt$delta_CFI[2] >= -0.004 & 
         tabla_completa_cnt$delta_RMSEA[2] <= 0.050, "Yes", "No"),
  # Scalar: ΔCFI ≥ -0.004 Y ΔRMSEA ≤ 0.010
  ifelse(tabla_completa_cnt$delta_CFI[3] >= -0.004 & 
         tabla_completa_cnt$delta_RMSEA[3] <= 0.010, "Yes", "No")
)
```

## Gender invariance

### Model fitting
```{r}
#| warning: false

config_sex <- cfa(model = model_syntax_3, 
              data = pisa,
              estimator = "WLSMV",
              group = "sex",
              sampling.weights = "SENWT",
              sampling.weights.normalization = "group",
              missing = 'pairwise',
              ordered = T)

metric_sex <- cfa(model = model_syntax_3, 
              data = pisa,
              estimator = "WLSMV",
              group = "sex",
              sampling.weights = "SENWT",
              sampling.weights.normalization = "group",
              ordered = T,
              missing = 'pairwise',
              group.equal= "loadings")

scalar_sex <- cfa(model = model_syntax_3, 
              data = pisa,
              estimator = "WLSMV",
              group = "sex",
              sampling.weights = "SENWT",
              sampling.weights.normalization = "group",
              ordered = T,
              missing = 'pairwise',
              group.equal= c("loadings", "intercepts"))

```

### Fit measures stracting

```{r}
fitmeasures_config_sex <- fitMeasures(config_sex,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

fitmeasures_metric_sex <- fitMeasures(metric_sex,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))

fitmeasures_scalar_sex <- fitMeasures(scalar_sex,
                         c("chisq", "df", "pvalue", "cfi", "tli", "rmsea"))
```

### Table processing
```{r}

tabla_completa_sex <- data.frame(
  Type = c("Gender", "", ""),
  Model = c("1. Configural", "2. Metric", "3. Scalar"),
  chisq = c(fitmeasures_config_sex["chisq"], fitmeasures_metric_sex["chisq"], fitmeasures_scalar_sex["chisq"]),
  df = c(fitmeasures_config_sex["df"], fitmeasures_metric_sex["df"], fitmeasures_scalar_sex["df"]),
  CFI = c(fitmeasures_config_sex["cfi"], fitmeasures_metric_sex["cfi"], fitmeasures_scalar_sex["cfi"]),
  TLI = c(fitmeasures_config_sex["tli"], fitmeasures_metric_sex["tli"], fitmeasures_scalar_sex["tli"]),
  RMSEA = c(fitmeasures_config_sex["rmsea"], fitmeasures_metric_sex["rmsea"], fitmeasures_scalar_sex["rmsea"])
)

# Deltas

tabla_completa_sex$delta_chisq <- c(NA, diff(tabla_completa_sex$chisq))
tabla_completa_sex$delta_df    <- c(NA, diff(tabla_completa_sex$df))
tabla_completa_sex$delta_CFI   <- c(NA, diff(tabla_completa_sex$CFI))
tabla_completa_sex$delta_RMSEA <- c(NA, diff(tabla_completa_sex$RMSEA))

# Deltas display

tabla_completa_sex$delta_CFI_display <- sapply(1:nrow(tabla_completa_sex), function(i) {
  if (i == 1) return(NA)
  
  if (tabla_completa_sex$delta_CFI[i] >= -0.004) {
    return("< -0.004")
  } else {
    return(sprintf("%.3f", tabla_completa_sex$delta_CFI[i]))
  }
})

tabla_completa_sex$delta_RMSEA_display <- sapply(1:nrow(tabla_completa_sex), function(i) {
  if (i == 1) return(NA)
  
  umbral <- if (i == 2) 0.050 else 0.010
  
  if (tabla_completa_sex$delta_RMSEA[i] <= umbral) {
    return(paste0("< ", umbral))
  } else {
    return(sprintf("%.3f", tabla_completa_sex$delta_RMSEA[i]))
  }
})

# ANOVA for significance

anova_sex <- anova(config_sex, metric_sex, scalar_sex)

tabla_completa_sex$p_value_delta_chisq <- case_when(anova_sex$`Pr(>Chisq)`< 0.01 ~ "< 0.01",
                                                  anova_sex$`Pr(>Chisq)`< 0.05 ~ "< 0.05",
                                                  anova_sex$`Pr(>Chisq)`> 0.05 ~ sprintf("%.2f", anova_sex$`Pr(>Chisq)`))
tabla_completa_sex$Decision <- c(
  NA,  # Configural
  # Metric: ΔCFI ≥ -0.004 Y ΔRMSEA ≤ 0.050
  ifelse(tabla_completa_sex$delta_CFI[2] >= -0.004 & 
         tabla_completa_sex$delta_RMSEA[2] <= 0.050, "Yes", "No"),
  # Scalar: ΔCFI ≥ -0.004 Y ΔRMSEA ≤ 0.010
  ifelse(tabla_completa_sex$delta_CFI[3] >= -0.004 & 
         tabla_completa_sex$delta_RMSEA[3] <= 0.010, "Yes", "No")
)
```

## Final table formatting
```{r}
tabla_completa_cnt_export <- tabla_completa_cnt %>%
  select(Type, Model, chisq, df, CFI, TLI, RMSEA,
         delta_chisq, delta_df, 
         delta_CFI_display, delta_RMSEA_display,
         p_value_delta_chisq, Decision) %>%
  rename(delta_CFI = delta_CFI_display,
         delta_RMSEA = delta_RMSEA_display)

tabla_completa_sex_export <- tabla_completa_sex %>%
  select(Type, Model, chisq, df, CFI, TLI, RMSEA,
         delta_chisq, delta_df, 
         delta_CFI_display, delta_RMSEA_display,
         p_value_delta_chisq, Decision) %>%
  rename(delta_CFI = delta_CFI_display,
         delta_RMSEA = delta_RMSEA_display)

tabla_unificada_pisa <- bind_rows(tabla_completa_cnt_export, tabla_completa_sex_export)

```

```{r}

colnames(tabla_unificada_pisa) <- c(
  "Type",
  "Model", 
  "\u03C7\u00B2",      # χ²
  "df",
  "CFI",
  "TLI",
  "RMSEA",
  "\u0394\u03C7\u00B2", # Δχ²
  "\u0394df",          # Δdf
  "\u0394CFI",         # ΔCFI
  "\u0394RMSEA",       # ΔRMSEA
  "p",
  "Decision"
)

ft_pisa <- flextable(tabla_unificada_pisa) %>%
  colformat_double(j = c("\u03C7\u00B2", "df", "\u0394\u03C7\u00B2", "\u0394df"), digits = 0) %>%
  colformat_double(j = c("CFI", "TLI", "RMSEA", "\u0394CFI", "\u0394RMSEA", "p"), digits = 3) %>%  border_remove() %>%
  hline_top(border = officer::fp_border(width = 2), part = "all") %>%
  hline_bottom(border = officer::fp_border(width = 2), part = "all") %>%
  hline(i = 1, border = officer::fp_border(width = 1.5), part = "header") %>%
  bold(part = "header") %>%
  font(fontname = "Calibri", part = "all") %>%
  fontsize(size = 9, part = "body") %>%
  fontsize(size = 10, part = "header") %>%
  padding(padding = 3, part = "all") %>%
  align(align = "center", part = "header") %>%
  align(align = "left", part = "body", j = 1:2) %>%
  align(align = "center", part = "body", j = 3:ncol(tabla_unificada_pisa)) %>%
  autofit() %>%
  fit_to_width(max_width = 7.5)


# Which cols to color
delta_cfi_col <- which(colnames(tabla_unificada_pisa) == "\u0394CFI")
delta_rmsea_col <- which(colnames(tabla_unificada_pisa) == "\u0394RMSEA")
decision_col <- which(colnames(tabla_unificada_pisa) == "Decision")
# Which rows to color
rows_cfi_fail <- which(!grepl("<", tabla_unificada_pisa[[delta_cfi_col]]) & 
                       !is.na(tabla_unificada_pisa[[delta_cfi_col]]))
rows_rmsea_fail <- which(!grepl("<", tabla_unificada_pisa[[delta_rmsea_col]]) & 
                         !is.na(tabla_unificada_pisa[[delta_rmsea_col]]))
rows_decision_no <- which(tabla_unificada_pisa$Decision == "No")
# Red when dont pass
if(length(rows_cfi_fail) > 0) {
  ft_pisa <- ft_pisa %>%
    bg(i = rows_cfi_fail, j = delta_cfi_col, bg = "#FFCCCB") %>%
    color(i = rows_cfi_fail, j = delta_cfi_col, color = "#8B0000")
}

if(length(rows_rmsea_fail) > 0) {
  ft_pisa <- ft_pisa %>%
    bg(i = rows_rmsea_fail, j = delta_rmsea_col, bg = "#FFCCCB") %>%
    color(i = rows_rmsea_fail, j = delta_rmsea_col, color = "#8B0000")
}

if(length(rows_decision_no) > 0) {
  ft_pisa <- ft_pisa %>%
    bg(i = rows_decision_no, j = decision_col, bg = "#FFCCCB") %>%
    color(i = rows_decision_no, j = decision_col, color = "#8B0000") %>%
    bold(i = rows_decision_no, j = decision_col)
}

ft_pisa
```

```{r}
#| eval: false

# Save tables
save_as_docx(ft_pisa, path = "output/figures/tables/invariance_table_pisa.docx")
save_as_image(ft_pisa, path = "output/figures/tables/invariance_table_pisa.png")
save_as_image(ft_pisa, path = "output/article/images/invariance_table_pisa.png")

```

# Descriptive analysis

## Latent means prediction

```{r}

fscores_df <- lavPredict(
  scalar_cnt,
  type = "lv", 
  method = "EBM",
  assemble = TRUE,
  append.data = TRUE
)

fscores_df |>
  summarise(
    total = n(),
    mean_gen_dse = mean(gen_dse, na.rm = TRUE),
    mean_spec_dse = mean(spec_dse, na.rm = TRUE)
  )

```

## Index and latent means correspondence analysis

### Index construction

```{r}
fscores_df <- fscores_df |> 
  mutate(
    gen_dse_index = rowMeans(pick(search_info, evaluate_info, collect_data, 
                                  explain_content, share_content, pair_collab, 
                                  write_text, create_media, change_settings), 
                             na.rm = TRUE),
                             
    spec_dse_index = rowMeans(pick(programming, identify_error, logical_solution), 
                              na.rm = TRUE)
  )
```

### Correlation index vs. latent mean

```{r}
cor_gen <- cor(fscores_df$gen_dse_index, fscores_df$gen_dse, 
               use = "complete.obs")
cor_spec <- cor(fscores_df$spec_dse_index, fscores_df$spec_dse, 
                use = "complete.obs")

n_global <- sum(!is.na(fscores_df$gen_dse))

global_row <- tibble(
  CNT = "GLOBAL", 
  n = n_global,
  cor_gen = cor_gen,
  cor_spec = cor_spec
)

cnt_cors <- fscores_df |>
  filter(!is.na(gen_dse)) |>
  group_by(CNT) |>
  summarise(
    n = n(),
    cor_gen = cor(gen_dse_index, gen_dse, use = "complete.obs"),
    cor_spec = cor(spec_dse_index, spec_dse, use = "complete.obs")
  )

final_table <- bind_rows(global_row, cnt_cors)

knitr::kable(final_table, digits = 3, caption = "Index and latent means correlations")
```


## Visualizations

### Index construction
```{r}
pisa <- pisa |> 
  mutate(
    gen_dse_index = rowMeans(pick(search_info, evaluate_info, collect_data, 
                                  explain_content, share_content, pair_collab, 
                                  write_text, create_media, change_settings), 
                             na.rm = TRUE),
                             
    spec_dse_index = rowMeans(pick(programming, identify_error, logical_solution), 
                              na.rm = TRUE)
  )
```

### Descriptives by gender

```{r}
pisa |> 
  filter(sex %in% c("Female", "Male")) |> 
  mutate(sex = as_factor(sex)) |> 
  droplevels() |> 
  table1(~ gen_dse_index + spec_dse_index | sex, 
         data = _,
         caption = "General and Specialized DSE scales descriptives by gender")

```

### Country comparison

```{r}

# Create data with country means
beeswarm_plot <- pisa |>
  group_by(CNT)|>
  summarise(mean_geneff = mean(gen_dse_index, na.rm = T),
            mean_speceff = mean(spec_dse_index, na.rm = T))|>
  ungroup()|>
  pivot_longer(cols = 2:3,
               names_to = "type",
               values_to = "score")

# Sort and extract labels
beeswarm_labels <- 
  rbind(
  beeswarm_plot%>%
  group_by(type) %>%
  arrange(score) %>%
  slice(c(1:3, (n() - 2):n())) %>%
  ungroup() %>%
  mutate(CNT = to_label(CNT)),
  
  beeswarm_plot%>%
    filter(CNT %in% c(""))%>%
    mutate(CNT = to_label(CNT))%>%
    slice(-3)
  )

# Plot
pisa_cnt_plot <- ggplot(beeswarm_plot, aes(y = score, x = type, color = type)) +
  geom_beeswarm(size = 2) +
  geom_text_repel(
    data = beeswarm_labels,
    aes(label = CNT),
    size = 3,
    color = "black",
    box.padding = 1.0,        
    point.padding = 0.1,      
    min.segment.length = 0,   
    max.overlaps = Inf,       
    force = 3,                
    force_pull = 0.3,         
    direction = "both",       
    seed = 999              
  ) +
  scale_color_manual(values = c("#5f5758","#fe3057")) +
  theme_minimal() +
  theme(
    legend.title = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    legend.position = "none",
    plot.caption = element_text(size = 15),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
  ) +
  scale_x_discrete(labels = c("General DSE", "Specialized DSE")) +
  scale_y_continuous(breaks = c(2,2.5,3,3.5,4), limits = c(2,4))

pisa_cnt_plot
```

### Gender gap country comparison

```{r}
# Create data in long format with gender means
gender_plot <- pisa |>  
  group_by(CNT, sex) |>  
  summarise(
    "General DSE" = mean(gen_dse_index, na.rm = TRUE),
    "Specialized DSE" = mean(spec_dse_index, na.rm = TRUE)
  ) |>  
  pivot_longer(
    cols = 3:4,
    names_to = "type",
    values_to = "score"
  ) |>  
  ungroup() |>  
  mutate(sex = to_label(sex)) |>  
  drop_na()

# Extract labels
country_labels_map <- gender_plot %>%
  distinct(CNT) %>%
  mutate(
    CNT_code = as.character(CNT),
    CNT_label = to_label(CNT)
  ) %>%
  select(CNT_code, CNT_label)

# Reorder the data by score for each type
gender_plot <- gender_plot %>%
  mutate(CNT_code = as.character(CNT)) %>%
  left_join(country_labels_map, by = "CNT_code") %>%
  group_by(type) %>%
  arrange(score) %>%
  mutate(CNT_label = factor(CNT_label, levels = unique(CNT_label))) %>%
  ungroup()


pisa_gender_plot <- ggplot(gender_plot, aes(y = CNT_label, x = score)) + 
  facet_grid(~type, scales = "free_y") +
  geom_line(aes(group = CNT_label), size = 0.5) +  
  geom_point(aes(color = sex, shape = sex), size = 2) +  
  scale_color_manual(
    name = "Gender",
    labels = c("Female", "Male"),
    values = c("Female" = "#FE3057", "Male" = "#5f5758")) +
  scale_shape_manual(
    name = "Gender",
    labels = c("Female", "Male"),
    values = c("Female" = 16, "Male" = 17)) + 
  theme_minimal() +  
  theme(
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    legend.position = "top",
    axis.text.y = element_text(size = 8, color = "black"),
    plot.margin = margin(10, 10, 10, 10),
    panel.spacing = unit(1.5, "lines"),
    text = element_text(size = 12),
    strip.text = element_text(size = 12),
    legend.text = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
  ) +
  coord_cartesian(clip = "off") +
  scale_x_continuous(breaks = c(2, 2.5, 3, 3.5, 4), limits = c(2, 4))

pisa_gender_plot
```

## Loadings and chi analysis (appendix)

```{r}

all_params <- parameterEstimates(scalar_cnt, standardized = TRUE)

factor_loadings_country <- all_params %>%
  filter(op == "=~")

test_stats <- lavInspect(scalar_cnt, "test")
chisq_values <- test_stats$scaled.shifted$stat.group
country_names_lavaan <- attr(test_stats, "info")$group.label

chisq_map <- data.frame(
  CNT = country_names_lavaan,
  Chisqr = as.numeric(chisq_values)
)

label_dictionary <- pisa %>%
  select(CNT) %>%
  distinct() %>%
  mutate(
    label_text = as.character(haven::as_factor(CNT)) 
  )

groups_code <- lavInspect(scalar_cnt, "group.label")

matched_labels <- label_dictionary$label_text[match(groups_code, label_dictionary$CNT)]

country_map <- data.frame(
  group = 1:length(groups_code),
  CNT = groups_code,
  Country = matched_labels
)

factor_loadings_final <- factor_loadings_country %>%
  left_join(country_map, by = "group") %>%
  left_join(chisq_map, by = "CNT") %>%
  select(Country, Item = rhs, Factor = lhs, Loading = std.all, Chisqr)


```

## Table for report

```{r}

datatable(
  factor_loadings_final, 
  rownames = FALSE,
  filter = 'top',
  options = list(
    pageLength = 15,
    autoWidth = TRUE
  ),
  caption = 'Tabla Interactiva: Cargas Factoriales y Desajuste (Chi2) por País'
) %>%
  formatRound(columns = 'Loading', digits = 3) %>% 
  formatRound(columns = 'Chisqr', digits = 2)

```

# Figures export

## Function for HD saving

```{r}
export_plot <- function(plot_object, 
                       filename,
                       format = c("png", "pdf", "tiff", "eps", "svg"),
                       width = 10,
                       height = 7,
                       dpi = 300,
                       units = "in",
                       bg = "white",
                       path = "./output/figures/") {
  
  format <- match.arg(format)
  
  full_filename <- paste0(filename, ".", format)
  full_path <- file.path(path, full_filename)
  

  ggsave(
    filename = full_path,
    plot = plot_object,
    width = width,
    height = height,
    dpi = dpi,
    units = units,
    bg = bg,
    device = format
  )
  
  message(paste0("✓ Gráfico exportado: ", full_path))
}
```

## Function for multiformat HD saving

```{r}
export_plot_multi <- function(plot_object,
                              filename,
                              formats = c("png", "pdf", "tiff"),
                              width = 10,
                              height = 7,
                              dpi = 300,
                              units = "in",
                              bg = "white",
                              path = "./output/figures/") {
  
  purrr::walk(formats, ~export_plot(
    plot_object = plot_object,
    filename = filename,
    format = .x,
    width = width,
    height = height,
    dpi = dpi,
    units = units,
    bg = bg,
    path = path
  ))
}
```


## Saving the plots in multiple formats

```{r}
#| eval: false

export_plot_multi(pisa_cnt_plot, "pisa_country_dse_comparison", 
                  formats = c("png", "pdf", "tiff"), width = 10, height = 6, dpi = 600, path = "./output/figures/plots/")

export_plot_multi(pisa_gender_plot, "pisa_gender_dse_gaps", 
                  formats = c("png", "pdf", "tiff"), width = 10, height = 6, dpi = 600, path = "./output/figures/plots/")
# Article
export_plot_multi(pisa_cnt_plot, "pisa_country_dse_comparison", 
                  formats = c("png"), width = 10, height = 6, dpi = 600, path = "./output/article/images/")

export_plot_multi(pisa_gender_plot, "pisa_gender_dse_gaps", 
                  formats = c("png"), width = 10, height = 6, dpi = 600, path = "./output/article/images/")

```